### 2025-03-23 src/main.tex
init

\documentclass[sigconf,natbib=false]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[ACM REP'25]{2025 ACM Conference on Reproducibility and Replicability}{July 29-31, 2025}{Vancouver, Canada}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmISBN{978-1-4503-XXXX-X/18/06}

\usepackage{tcolorbox}
\tcbuselibrary{theorems}
\usepackage{cleveref}
\newtcbtheorem[]{trap}{Pitfall}{colback=black!5,colframe=black!35,fonttitle=\bfseries}{th}
\newtcbtheorem[]{lesson}{Takeaway}{colback=black!5,colframe=black!35,fonttitle=\bfseries}{th}
\newcommand{\repro}{reproducibility}
\newcommand{\Repro}{Reproducibility}
\newcommand{\transpo}{\emph{Transposition}}
\newcommand{\flavour}{\emph{flavour}}
\newcommand{\flavours}{\emph{flavours}}
\newcommand{\ie}{\emph{i.e.,}}
\newcommand{\eg}{\emph{e.g.,}}
\newcommand{\nix}{\emph{Nix}}
\newcommand{\nixos}{\emph{NixOS}}
\newcommand{\nxc}{\emph{NixOS Compose}}
\newcommand{\enos}{\emph{EnOSlib}}
\newcommand{\grid}{\emph{Grid'5000}}
\newcommand{\kam}{\emph{Kameleon}}
\newcommand{\kad}{\emph{Kadeploy}}
\newcommand{\mel}{\emph{Melissa}}
\newcommand{\store}{\emph{Nix Store}}
\newcommand{\ad}{Artifact Description}
\newcommand{\aeval}{Artifact Evaluation}
\newcommand{\adae}{\ad/\aeval}
\newcommand{\df}{\texttt{Dockerfile}}
\newcommand{\ecg}{\texttt{ecg}}
\newcommand{\todo}[1]{{\color{red}{TODO: #1}}}
\usepackage{hyperref}
\usepackage{array}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{siunitx}
%\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{calc}
%\usepackage{tabularx}
\usepackage{fontawesome}
\usepackage[para,online,flushleft]{threeparttable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{amsmath,amsfonts}
\usepackage{textcomp}
\usepackage[inline]{enumitem}  % enhanced enumerations
\usepackage{listings}
\definecolor{commentcolour}{rgb}{0.04,0.43,0.17}
\definecolor{keywordcolour}{rgb}{0.65,0.15,0.64}
\definecolor{backcolour}{rgb}{1,1,1}
\definecolor{linenumbercolour}{rgb}{0.1,0.1,0.1}
\definecolor{stringcolour}{rgb}{0.56,0.06,0.49}
\lstdefinestyle{mystyle}{
  backgroundcolor=\color{backcolour},   
  commentstyle=\color{commentcolour},
  keywordstyle=\color{keywordcolour},
  numberstyle=\tiny\color{linenumbercolour},
  stringstyle=\color{stringcolour},
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=false,         
  breaklines=true,                 
  captionpos=b,                    
  keepspaces=true,                 
  numbers=left,                    
  numbersep=5pt,                  
  showspaces=false,                
  showstringspaces=false,
  showtabs=false,                  
  tabsize=2,
  framexrightmargin=-12pt
}
\lstset{style=mystyle, frame=single}

%\usepackage[backend=biber,style=trad-abbrv,firstinits=true]{biblatex}
\usepackage[
  datamodel=software,
  style=trad-abbrv,
  backend=biber
]{biblatex}
\addbibresource{references.bib}
\usepackage{software-biblatex}

\begin{document}

\title{%
  % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
  Longitudinal Study of the Software Environment Produced by Dockerfiles from Research Artifacts: Initial Design and Results
}

\author{Quentin Guilloteau}
\email{quentin.guilloteau@inria.fr}
\affiliation{%
  \institution{INRIA}
  \country{France}
}

% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  The reproducibility crisis struck all the fields of Science, including Computer Science.
  To tackle the issue, the CS community set up artifact evaluation processes in conferences and journals to evaluate the reproducibility of the results shared in the publications.
  Authors thus share their artifacts included code, data and software environment required to reproduce the results.
  One method for sharing the software environment, that has been ``advertised'' by conferences and journals, is to use container technologies such a Docker and Apptainer.
  However, these tools relies on non-reproducible tools, hence producing non-reproducibile containers.

  In this paper, we present a tool and methodology to evaluate the variation in software environments through time of container images from research artifacts.
  We also present initial results on a small set of \df s from the EuroPar'24 conference.
\end{abstract}

%% %% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Reproducibility, Artifact Evaluation, Badges, Lifetime}

\received{12 February 2024}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

%% PAPER STARTS HERE -----------------------------------------------------------------------------------------------

\section{Introduction}

% \todo{split into intro and related work?}
% 
% \begin{itemize}
% \item Reproducibility crisis \cite{baker500ScientistsLift2016}
% \item Also in computer science \cite{collberg_repeatability_2015}, but most impact is from the missing environment.
% \item Artifact evaluation in conferences \cite{kidwell2016badges}
% \item Software environment is a problem/difficul aspect \cite{mytkowicz_producing_nodate}
% \item conferences recommand to use containers to share the software environmnent,
% \item more and more papers are linking code \cite{paperswithcode, kang2023papers, hong2013software}
% \item a lot of papers proposing solution around virtualisation and container \cite{brammer2011paper, brinckman2019computing}
% \item but container recipes are based on non-reproducibile tools (package managers and bad practices) \cite{guilloteau2024longevity, guilloteau2024frustrations}
% \item even though Nix \cite{dolstra_nix_2004} and Guix \cite{courtes_functional_2013} already very good solutions but not used enough \cite{guilloteau2024longevity}
% \item trade-off between storing the result of the image builts by the authors and the recipe \cite{software_heritage_2017, zenodo, figshare}
% \item we need the dev environment to \emph{inspect} and introduce \emph{variation} \cite{mercier2018considering}. \cite{blinry, nvidia_cuda_lifetime}
% \end{itemize}

The entire scientific community has been struck by the ``Reproducibility Crisis''~\cite{baker500ScientistsLift2016}, and Computer Science does not make exception~\cite{collberg_repeatability_2015}.
In order to improve the reproducibility of articles in Computer Science, the community set up \adae\ processes~\cite{kidwell2016badges} where authors would submit (if they want) the artifact required to reproduced the main results from their article.
Then a set of artifact reviewers would try to reproduce those main results given the artifact from the authors.
Based on the result of the reproduction attempt, the article is then awarded \emph{badges} to promote its reproducibility aspect.
The link to the artifacts can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, experiments workflow) \cite{paperswithcode, kang2023papers, hong2013software}.

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible varability of results given a badly captured software environment \cite{mytkowicz_producing_nodate} \todo{more}.
To tackle this issue, the conference \ae\ committees have been pushing the authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment.
However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
Moreover, in the case of usual virtualisation techiniques to generate containers or virtual machines, users have to write the recipe made of a list of commands to execute from a base image.
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
These base images can also be removed by the administrators making all the recipes unbuildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results, and thus ``breaking'' the reproducibility of the results.
One solution would be to store the container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
But this come with the downside of requiring more storage space than just storing the recipe (binary vs. text file).
This practice has thus some \textbf{severe scalability and sustainability issues in terms of longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

\todo{this is already some kind of conlusion...}
A better solution would be to use tools such as Nix~\cite{dolstra_nix_2004} or Guix~\cite{courtes_functional_2013} which are focused on reproducibility and which are able to rebuild the \emph{exact same} software environment from a textual description and with longevity properties~\cite{courtes2024source}, but these tools are unfortunately poorly adopted by the community~\cite{guilloteau2024longevity}.

In this paper, we propose our \textbf{inital design of a longitudinal study of the variations through time of the software environment produced by \df s from research artifacts.}
We show initial results from five \df s extracted from the EuroPar'24 conference reproducibility artifacts built every month for six months.
\todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}

% We aim to answer the following questions:
% 
% \begin{enumerate}
% \item How does the packages evolve through time when captured in a \texttt{Dockerfile}?
% \item Which packaging tool is the most victim to variation in its resulting software environment?
% \item Which are the most volatile packages?
% \item Which practices in \df s are the most damaging to reproducibility?
% \end{enumerate}

The following of this paper is organized as follows:
Section~\ref{sec:ecg} presents our software framework for this longitudinal study.
The procol is presented in Section~\ref{sec:protocol}.
Section~\ref{sec:results} presents the initial results of a small scale version of the study.
Finally, we conclude and give perspectives in Section~\ref{sec:conclusion}.

% \newpage
\section{The \ecg\ Framework}\label{sec:ecg}

%% \begin{itemize}
%% \item `ecg`, Python tool, Licence, lines of code
%% \item architecture (nickel, what is supported)
%% \item process (submit new artifacts)
%% \item end goal of the project
%% \end{itemize}

In this section, we present \ecg, a tool for the periodic building of \df s from research artifacts and the collection of software environment informations.
\ecg\ is a 280-line Python3 script under the GPLv3 licence.

\subsection{Artifact representation}\label{sec:nickel}

An \df\ from an artifact is represented as a Nickel~\cite{nickel} file.
Nickel is a configuration language with functional paradigm.
One of it's feature is the possibility go ``verify'' that an configuration is ``well-formed'' with a \emph{contract}.
This Nickel file is then \emph{verified} against the contract (\todo{link to contract}) and translated to a JSON file fed to \ecg.

\begin{lstlisting}[caption=Example of Artifact representation in the Nickel format.]
{
  version = "1.0",
  artifact_url = "https://zenodo.org/.../artifact.zip",
  type = "zip",
  doi = "10.5281/zenodo.XXXXXXXX",
  conf_date = 2024,
  virtualization = "docker",
  buildfile_dir = "artifact",
  package_managers = [ "dpkg", "pip" ]
  git_packages = [
    { name = "spack", location = "/home/user/spack" }
  ],
  misc_packages = [
    {
      name = "cmake-3.22.2-linux",
      url = "https://github.com/Kitware/CMake/releases/download/v3.22.2/cmake-3.22.2-linux-x86_64.sh"
    }
  ]
}
\end{lstlisting}

\subsection{What do we capture?}

\subsubsection{Hash of the Artifact}

The articles give the link to the artifact.
This artifact can be shared in several ways, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
Hence, we capture the result of the download at the given link (success or error).
In case of a succesful download, we also compute the cryptographic hash of the result of download to capture any variation in the source of the artifact which could explain later variations in the resulting software environment.

For instance, authors can give the link to a Zenodo archive that will \emph{always} point to the latest version of an artifact.
Hence, the content of the artifact behind this link can change through time.

\subsubsection{Docker Building Error}

In the case of a \df s failing to build, we capture the error (missing image, failing command, etc.).
In fact, we are only able to capture the first error of the build even if there might be several errors in the building process.
The Docker error is extracting by \texttt{grep}ing the building process log.

\subsubsection{List of Installed Packages and their Version}

The software environment can be composed of several packages.
Theses packages can have been installed by different means.
The most common way is through package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
But there are also installed packages for programming languages such as with \texttt{pip} for Python.
Another way to install package is to download the sources and build them locally.
\ecg\ is able to capture these different ways to install packages.
Hence, for each package, we extract the package name, the version (the \texttt{git} commit if using \texttt{git}), and the tool used to install the package.

As of for now, \ecg\ is able to query information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are install by source (\texttt{package\_manager} in Listing 1).
In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing 1), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container.
This allows us to trace the version of the package.
In the case of a simple downloading of the sources of a package (\texttt{misc\_packages} in Listing 1), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being dowloaded.
The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
Hence, \ecg\ will download \emph{localy} the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.

\subsection{Regularly Building Containers}

Once the artifacts have been described in the Nickel representation, we are able to build the \df s associated.
We plan to build all the artifacts at the start of each month.
The goal is to build the artifacts for one year.

\todo{building patterns, exponential decay}

With the total computing and energetical costs in mind, we limit the building times of each container to 10 minutes.

\subsection{Workflow}\label{sec:workflow}

\begin{figure*}
  \centering
  \fbox{
  \includegraphics[width=\textwidth]{./figs/workflow.pdf}
  }
  \caption{
    Workflow of \ecg.
    Each description of an artifact is verified with the Nickel contract and then converted in a JSON representation.
    This JSON representation is then read by \ecg\ to
    \textit{(i)} download the artifact,
    \textit{(ii)} compute the hash of its content,
    \textit{(iii)} build the container from the \df, and
    \textit{(iv)} extract the software environment information from the built container.
    \ecg\ outputs files containing the information about the artifact and its \df\ .
  }
  \label{fig:workflow}
\end{figure*}

The workflow of ecg is depicted in Figure~\ref{fig:workflow}, and is executed on machines of the \grid\ platform \cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
Theses machines are x86 machines.
We make sure to use an empty local Docker cache before building the containers.

The workflow itself is ruled by a Snakemake~\cite{koster2012snakemake} description, where the software environment is managed with Nix~\cite{dolstra_nix_2004}.

\section{Protocol}\label{sec:protocol}

In this section, we present the initial protocole of our longitudinal study.

For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
From the artifact and the \df, we write the associated Nickel file: where is stored the artifact? how to extract the \df\ out of the artifact? which package managers are used?
Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
We run \ecg\ the closer to the start of each month as possible.
\ecg\ is executed on the (x86) machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
The local Docker cache of these machine is cleaned before the building of each container.
Each container has an allocated time of 10 minutes to build.

We will execute \ecg\ on each artifact every month for a full year (13 months).

\paragraph{Contributions}

This protocol leaves a lot of room for contributions on the writing of the Nickel descriptions files.

% \todo{}
% 
% \begin{itemize}
% \item we convert the dockerfiles to nickel description
% \item periodic rebuild of the dockerfiles
% \item g5k
% \item empty cache
% \item max 10 mins
% \item contributions
% \end{itemize}

% \newpage
\section{Preliminary Results}\label{sec:results}

This section presents the information we would like to extract from the data gathered.

% And there is no ``warning message'' that the environment is different.

We wrote the description in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a container to share their software environments, and built the associated \df s once every month.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, and March 12$^{th}$ 2025.
The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
%\todo{when was the artifact submission, and publicaiton of the papers for europar?}

\subsection{Hash of Artifacts}

No variation of the artifact sources, via the computation of their cryptographic hash, was observed.

\subsection{Build Status}

All the artifacts managed to successfully produce a container.

\subsection{Software Environment}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{./figs/results_per_artifact.pdf}
  \caption{
    Evolution of the packages in the software environment of each container through time.
    Each container has been rebuilt once a month.
    The color of the bar corresponds to the month when a specific version of a package has been introduced in the software environment.
    We can see that the proportion of package versions similar to the versions in the initial build is decreasing throught the months.
  }
  \label{fig:results_per_artifact}
\end{figure*}

Figure~\ref{fig:results_per_artifact} depicts the evolution through time of the software environment produced by the \df\ of each studied artifact.
We can see that even after \emph{a single month} of the initial build, the software environment is already different for \emph{all} of the studied artifacts.
This is a severe threat to reproducibility as one month is the time-frame for a artifact evaluation process in conferences.
This means that the \df\ submitted by the authors can generate different software environment between the submission of the artifact and the reproduction attempt of the reviewers.

%\todo{a word on from ubuntu:latest and apt-get update}.

We can see that every month, new versions of packages are being introduced in the software environments of the containers.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{./figs/results_per_tool.pdf}
  \caption{
    plop
  }\label{fig:results_per_tool}
\end{figure}

Figure~\ref{fig:results_per_tool} depicts the evolution of the package version managed by each tool: \texttt{dpkg}, \texttt{git}, \texttt{pip} and manual download and build (\texttt{misc}).
We can see that the tools introducing variation in the software environment are \texttt{dpkg} and \texttt{pip}.

The packages that changed the most in terms of versions are:
\begin{enumerate}
\item \texttt{linux-libc-dev:amd64} (installed via \texttt{dpkg}) with 6 different versions for each artifact (1 per new build attempt)
\item \texttt{fonttools} (installed via \texttt{pip}) with 5 different versions for the artifacts \texttt{geijer\_how} and \texttt{wolff\_fast}
\item \texttt{numpy} (installed via \texttt{pip}) with 5 different versions for the artifact \texttt{wolff\_fast}
\end{enumerate}

% \newpage
\section{Conclusion and Perspectives}\label{sec:conclusion}

% \subsection{Conclusion}
% 
% In this paper, we presented an initial version of \ecg, and its workflow, a tool to study the variations through time of containers from research artifacts.
% We evaluated the evolution of the software environment produced by \df\ from five artifacts from the EuroPar'24 conference.
% We observe that, even at this small scale, after only one month, there were variations in the produced software environments.
% Out of the popular tools to manage the software environment in a container, we have seen that \texttt{dpkg} and \texttt{pip} were the ones that introducing more undesired variations.
% 
% \df s are not solutions to capture on the \emph{long term} the software environment.
% There is a trade-off between storing the Docker images built by the authors and their recipe.
% Use Nix/Guix + Software Heritage.
% 
% Even if we only looked at \df s, this study could be extended to other technologies (Apptainer, virtual machines), but it was shown in \cite{guilloteau2024longevity} that \df s was the main solution to provided virtualized software environments.
% 
% In this paper, we presented the initial design of our framework to evaluate the longevity of the software environments generated with \df s from research artifacts.
% 
% We propose three possible scenarios for the longevity of the sharing of software environment:
% 
% \paragraph{Scenario 1: ``Business as usual''}
% Authors keep submitting non-reproducible \df s and conferences keep accepting them
% 
% \paragraph{Scenario 2: Storing everything forever}
% Authors store their version of the built container onto long-term storage (\eg\ Zenodo, Figshare), and artifact reviewers do not try to rebuild the \df\ but simply download the container.
% While this solution would allow the artifact reviewers and the future researchers to use the exact same software environment as the authors, it also raises \emph{sustainability questions}.
% Indeed, with the increase of scientific publications and the increase in artifact sharing in publications, can we community continue to store the built containers which ``weight'' orders of magnitude more than the recipe to generate them?
% Services such as Zenodo and Figshare are free, and backed up by public funding.
% Can these services run out of storage? or ask for a financial participation to archive data?
% 
% 
% \paragraph{Scenario 3: Change of practice}
% In this scenario, the community, \ie\ the conferences and the authors, change their practices related to reproducibility and software environment sharing.
% The community starts adopting solutions made for reproducibility of software environment (\ie\ Nix and Guix).
% Authors then simply have to share the recipe (text files) of their software environment for the artifact reviewers and future researchers to reproduce the results of their article.
% Moreover, Nix and Guix allow for the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.
% 
% 
% \subsection{Perspectives}
% 
% This paper aims to be an initial version of the design of a future larger study around the longevity of containers.
% We will thus write the descriptions of more artifact from more conferences, aiming for around a hundreds total containers, from conferences of various fields of Computer Science with various ``technical'' levels.
% 
% This initial work also raised more questions:
% \begin{itemize}
% \item Does the reduction of the proportion of package versions similar to the initial build will continue decreasing or will it eventually reach a ``stable state''? (for example 50\%)
% \item ...
% \end{itemize}

% -----------------------------------------------------------------

This paper presented the initial design of our longitudinal study of the evolution of the software environments produced by \df s.
We presented in Section~\ref{sec:ecg} \ecg, and its workflow (Section~\ref{sec:workflow}), a tool to download, build and extract information from research artifacts \df.
Section~\ref{sec:protocol} presented the protocol to answer our scientific questions.
We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every months for six months.
This initial study releaved that the software environments produced by the considered \df s variate \emph{every month} which highlights the lack of reproducibility of such practices.

A practice to deal with the lack of reproducibility of the \df s is for the authors to store the built container onto some long-time storage such as Zenodo.
However, this practice raises serious questions in terms of scalability and sustainability in terms of storage.
The lack of reproducibility of \df s shoud not be hidden with long-term storage: the practices from the conferences, journals, and the authors need to change.

%
% /////////////////////////////////////////////////////////////////////////////////////////////////////

\section*{Acknowledgments}

DAPHNE, Antoine?

Experiments presented in this paper were carried out using the Grid'5000 testbed, supported by a scientific interest group hosted by Inria and including CNRS, RENATER and several Universities as well as other organizations (see https://www.grid5000.fr).

%% PAPER ENDS HERE -----------------------------------------------------------------------------------------------

\newpage
\newpage
%\bibliographystyle{sty/ACM-Reference-Format}
%\bibliography{references}
\printbibliography

\end{document}

### 2025-03-23 src/main.tex
[text] small changes

@@
   % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
-  Longitudinal Study of the Software Environment Produced by Dockerfiles from Research Artifacts: Initial Design and Results
+  Longitudinal Study of the Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
 }
@@
 \author{Quentin Guilloteau}
-\email{quentin.guilloteau@inria.fr}
+\email{Quentin.Guilloteau@inria.fr}
 \affiliation{%
@@
 
-\subsection{What do we capture?}
+\subsection{What do we capture?}\label{sec:ecg:capture}
 
@@
 
-\subsection{Regularly Building Containers}
-
-Once the artifacts have been described in the Nickel representation, we are able to build the \df s associated.
-We plan to build all the artifacts at the start of each month.
-The goal is to build the artifacts for one year.
-
-\todo{building patterns, exponential decay}
-
-With the total computing and energetical costs in mind, we limit the building times of each container to 10 minutes.
-
 \subsection{Workflow}\label{sec:workflow}
@@
 
+\subsection{Contributions}
+
+This framework leaves a lot of room for contributions on the writing of the Nickel descriptions files.
+The community can contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.
+
 \section{Protocol}\label{sec:protocol}
@@
 
-For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
-From the artifact and the \df, we write the associated Nickel file: where is stored the artifact? how to extract the \df\ out of the artifact? which package managers are used?
-Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
-We run \ecg\ the closer to the start of each month as possible.
-\ecg\ is executed on the (x86) machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
-The local Docker cache of these machine is cleaned before the building of each container.
-Each container has an allocated time of 10 minutes to build.
+\begin{enumerate}
+\item For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
+\item From the artifact and the \df, we write the associated Nickel file: where is stored the artifact? how to extract the \df\ out of the artifact? which package managers are used?
+\item Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
+\item We run \ecg\ the closer to the start of each month as possible.
+\item The local Docker cache of these machine is cleaned before the building of each container.
+\item We will execute \ecg\ on each artifact every month for a full year (13 months).
+\end{enumerate}
 
-We will execute \ecg\ on each artifact every month for a full year (13 months).
+\paragraph{Details}
 
-\paragraph{Contributions}
-
-This protocol leaves a lot of room for contributions on the writing of the Nickel descriptions files.
+\begin{itemize}
+\item Each container has an allocated time of 10 minutes to build.
+\item \ecg\ is executed on the (x86) machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
+\end{itemize}
 
@@
 
-We wrote the description in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a container to share their software environments, and built the associated \df s once every month.
+We wrote the description in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and built the associated \df s once every month for six months.
 % "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
@@
 
+The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).
+
 \subsection{Hash of Artifacts}
+
+Four out the five articles provided a Zenodo archive to the artifact, the other provided a github link with a specified \texttt{git} tag.
 
@@
 \subsection{Software Environment}
+
+\todo{check which base image, tag of base image and if apt-get update}
 
@@
 This means that the \df\ submitted by the authors can generate different software environment between the submission of the artifact and the reproduction attempt of the reviewers.
-
-%\todo{a word on from ubuntu:latest and apt-get update}.
 
@@
 
-DAPHNE, Antoine?
+Antoine?
 
 Experiments presented in this paper were carried out using the Grid'5000 testbed, supported by a scientific interest group hosted by Inria and including CNRS, RENATER and several Universities as well as other organizations (see https://www.grid5000.fr).
+
+This research was funded, in whole or in part, by the European Union’s Horizon 2020 research and innovation programme under grant agreement No. 957407 as DAPHNE.
 

### 2025-03-23 src/main.tex
[text] fitting on 4 pages

@@
 In order to improve the reproducibility of articles in Computer Science, the community set up \adae\ processes~\cite{kidwell2016badges} where authors would submit (if they want) the artifact required to reproduced the main results from their article.
-Then a set of artifact reviewers would try to reproduce those main results given the artifact from the authors.
+Then, artifact reviewers would try to reproduce those main results given the artifact from the authors.
 Based on the result of the reproduction attempt, the article is then awarded \emph{badges} to promote its reproducibility aspect.
-The link to the artifacts can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, experiments workflow) \cite{paperswithcode, kang2023papers, hong2013software}.
+The link to the artifacts can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, workflow)~\cite{paperswithcode, kang2023papers, hong2013software}.
 
 One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
-Several works already shed light on the possible varability of results given a badly captured software environment \cite{mytkowicz_producing_nodate} \todo{more}.
+Several works already shed light on the possible varability of results given a badly captured software environment \cite{mytkowicz_producing_nodate}.
 To tackle this issue, the conference \ae\ committees have been pushing the authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment.
@@
 
-\todo{this is already some kind of conlusion...}
-A better solution would be to use tools such as Nix~\cite{dolstra_nix_2004} or Guix~\cite{courtes_functional_2013} which are focused on reproducibility and which are able to rebuild the \emph{exact same} software environment from a textual description and with longevity properties~\cite{courtes2024source}, but these tools are unfortunately poorly adopted by the community~\cite{guilloteau2024longevity}.
+%\todo{this is already some kind of conlusion...}
+%A better solution would be to use tools such as Nix~\cite{dolstra_nix_2004} or Guix~\cite{courtes_functional_2013} which are focused on reproducibility and which are able to rebuild the \emph{exact same} software environment from a textual description and with longevity properties~\cite{courtes2024source}, but these tools are unfortunately poorly adopted by the community~\cite{guilloteau2024longevity}.
 
@@
 We show initial results from five \df s extracted from the EuroPar'24 conference reproducibility artifacts built every month for six months.
-\todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}
+% \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}
 
@@
 Nickel is a configuration language with functional paradigm.
-One of it's feature is the possibility go ``verify'' that an configuration is ``well-formed'' with a \emph{contract}.
+One of it's feature is the possibility to ``verify'' that an configuration is ``well-formed'' with a \emph{contract}.
 This Nickel file is then \emph{verified} against the contract (\todo{link to contract}) and translated to a JSON file fed to \ecg.
+The contract is versionned to take into account the potential future changes to the descriptions and the contract.
 
-\begin{lstlisting}[caption=Example of Artifact representation in the Nickel format.]
+\begin{lstlisting}[caption=Example of Artifact Nickel representation]
 {
@@
 
-\subsection{What do we capture?}\label{sec:ecg:capture}
+\subsection{Information captured by \ecg}\label{sec:ecg:capture}
 
@@
 
-The workflow of ecg is depicted in Figure~\ref{fig:workflow}, and is executed on machines of the \grid\ platform \cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
-Theses machines are x86 machines.
+The workflow of \ecg\ is depicted in Figure~\ref{fig:workflow}.
 We make sure to use an empty local Docker cache before building the containers.
-
-The workflow itself is ruled by a Snakemake~\cite{koster2012snakemake} description, where the software environment is managed with Nix~\cite{dolstra_nix_2004}.
+The workflow itself is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment is managed with Nix~\cite{dolstra_nix_2004}.
 
@@
 
-\todo{check which base image, tag of base image and if apt-get update}
+\begin{table}
+  \centering
+  \scalebox{0.925}{
+  \begin{tabular}{llll}
+    \toprule
+    Artifact & Base image & Version & \texttt{update}?\\
+    \midrule
+    \texttt{canon\_solving} & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{geijer\_how}    & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{hiraga\_peanuts}& \texttt{devcontainers/cpp} & \texttt{1-debian-12} & Yes \\
+    \texttt{munoz\_fault}   & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{wolff\_fast}    & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \bottomrule
+  \end{tabular}
+  }
+  \caption{Information about the \df s from the study}\label{tab:info}
+\end{table}
+
+Table~\ref{tab:info} summarizes the information about the studied \df s.
+We can see that most of the container used the \texttt{ubuntu} base image and that \emph{all} the \df s specified the version of the base image to use, thus limiting the variations in terms of produced software environment.
+However, all of the \df s called the \texttt{apt-get update} command, which fetches every time the latest versions of the packages and thus does introduces variations.
 
@@
   \caption{
-    plop
+    Evolution through time of the packages in the software environments managed by each tool present in the studied artifacts.
   }\label{fig:results_per_tool}
@@
 We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every months for six months.
-This initial study releaved that the software environments produced by the considered \df s variate \emph{every month} which highlights the lack of reproducibility of such practices.
+
+While this initial study only considered five \df s, it releaved that \textbf{the software environments produced by the considered \df s variate \emph{every month}} which highlights the lack of reproducibility of such practices.
+The question of the significance of these results on such a small sample will need to be answered in a larger scale study.
 
@@
 The lack of reproducibility of \df s shoud not be hidden with long-term storage: the practices from the conferences, journals, and the authors need to change.
+
+\paragraph{Perspectives}
+
+This paper aimed to be an initial version of the design of a future larger study around the longevity of containers.
+We will consider not only Docker containers, but also other containerization tools such as Apptainer.
+We will thus write the descriptions of more artifacts from more conferences, aiming for around a hundreds total containers, from conferences of various fields of Science with various Computer Science technical levels.
 
@@
 
-Antoine?
+%Antoine?
 

### 2025-03-23 scripts/results_per_artifact.R
[R] forgot to commit the R scripts

@@ -1,0 +1,38 @@
+library(tidyverse)
+args = commandArgs(trailingOnly=TRUE)
+infile <- args[1]
+outfile <- args[2]
+
+df <- read_csv(infile, col_names = c("pkg", "version", "tool", "artifact", "date")) %>%
+	mutate(date = date(as_datetime(date)))
+
+dates <- df %>% pull(date) %>% unique()
+
+p <- df %>%
+	group_by(artifact, tool, pkg) %>%
+	arrange(date) %>%
+	nest() %>%
+	rowwise() %>%
+	mutate(data2 = list(data %>% arrange(date) %>% mutate(rank = .$date[match(version, .$version)]))) %>%
+	select(-data) %>%
+	unnest_wider(data2) %>%
+	unnest_longer(c("version", "date", "rank")) %>%
+	group_by(artifact, date) %>%
+	mutate(total_pkgs = n()) %>%
+	group_by(artifact, date, rank, total_pkgs) %>%
+	summarize(total = n()) %>%
+	mutate(rank = factor(rank)) %>%
+	mutate(rank = factor(rank, levels=rev(levels(rank)))) %>%
+	ggplot(aes(x = factor(date), y = 100 * total / total_pkgs, fill = rank)) +
+	geom_col() +
+	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 5, label = total), angle=0, size=2.0) +
+	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 10, label = paste("(",round(100*total/total_pkgs, digits = 1), "%)",sep="")), angle=0, size=1.5) +
+	ylab("Packages in environment [%]") +
+	facet_wrap(~artifact, ncol = 5) +
+	ggtitle("Evolution of the packages versions through time") +
+	scale_fill_grey("Month when the package version was introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
+	scale_x_discrete("Months after initial build", labels=append(c("Initial"), seq(length(dates)-1))) +
+	theme_bw() +
+	theme(legend.position = "bottom", strip.background = element_blank())
+
+ggsave(plot=p,outfile, width=8, height=3.25)

### 2025-03-23 scripts/results_per_tool.R

@@ -1,0 +1,41 @@
+library(tidyverse)
+args = commandArgs(trailingOnly=TRUE)
+infile <- args[1]
+outfile <- args[2]
+
+df <- read_csv(infile, col_names = c("pkg", "version", "tool", "artifact", "date")) %>%
+	mutate(date = date(as_datetime(date)))
+
+dates <- df %>% pull(date) %>% unique()
+
+p <- df %>%
+	group_by(artifact, tool, pkg) %>%
+	arrange(date) %>%
+	nest() %>%
+	rowwise() %>%
+	mutate(data2 = list(data %>% arrange(date) %>% mutate(rank = .$date[match(version, .$version)]))) %>%
+	select(-data) %>%
+	unnest_wider(data2) %>%
+	unnest_longer(c("version", "date", "rank")) %>%
+	group_by(tool, date) %>%
+	mutate(total_pkgs = n()) %>%
+	group_by(tool, date, rank, total_pkgs) %>%
+	summarize(total = n()) %>%
+	mutate(rank = factor(rank)) %>%
+	mutate(rank = factor(rank, levels=rev(levels(rank)))) %>%
+	ggplot(aes(x = factor(date), y = 100 * total / total_pkgs, fill = rank)) +
+	geom_col() +
+	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 5, label = total), angle=0, size=2.0) +
+	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 15, label = paste("(",round(100*total/total_pkgs, digits = 1), "%)",sep="")), angle=0, size=1.5) +
+	ylab("Packages in environment [%]") +
+	facet_wrap(~tool, ncol = 2) +
+	ggtitle("Evolution of the packages versions through time") +
+	scale_fill_grey("Month when the package version\nwas introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
+	scale_x_discrete("Months after initial build", labels=append(c("Initial"), seq(length(dates)-1))) +
+	theme_bw() +
+        theme(legend.position = "bottom", strip.background = element_blank(),
+          plot.title.position = "plot",
+          plot.caption.position =  "plot",
+          legend.location = "plot")
+
+ggsave(plot=p,outfile, width=4, height=4.25)

### 2025-03-23 src/main.tex
aspell + check text

@@
   One method for sharing the software environment, that has been ``advertised'' by conferences and journals, is to use container technologies such a Docker and Apptainer.
-  However, these tools relies on non-reproducible tools, hence producing non-reproducibile containers.
+  However, these tools relies on non-reproducible tools, hence producing non-reproducible containers.
 
@@
 One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
-Several works already shed light on the possible varability of results given a badly captured software environment \cite{mytkowicz_producing_nodate}.
+Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate}.
 To tackle this issue, the conference \ae\ committees have been pushing the authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment.
 However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
-Moreover, in the case of usual virtualisation techiniques to generate containers or virtual machines, users have to write the recipe made of a list of commands to execute from a base image.
+Moreover, in the case of usual virtualization techniques to generate containers or virtual machines, users have to write the recipe made of a list of commands to execute from a base image.
 This approach has several limitations.
@@
 
-In this paper, we propose our \textbf{inital design of a longitudinal study of the variations through time of the software environment produced by \df s from research artifacts.}
+In this paper, we propose our \textbf{initial design of a longitudinal study of the variations through time of the software environment produced by \df s from research artifacts.}
 We show initial results from five \df s extracted from the EuroPar'24 conference reproducibility artifacts built every month for six months.
@@
 Section~\ref{sec:ecg} presents our software framework for this longitudinal study.
-The procol is presented in Section~\ref{sec:protocol}.
+The protocol is presented in Section~\ref{sec:protocol}.
 Section~\ref{sec:results} presents the initial results of a small scale version of the study.
@@
 
-In this section, we present \ecg, a tool for the periodic building of \df s from research artifacts and the collection of software environment informations.
-\ecg\ is a 280-line Python3 script under the GPLv3 licence.
+In this section, we present \ecg, a tool for the periodic building of \df s from research artifacts and the collection of software environment information.
+\ecg\ is a 280-line Python3 script under the GPLv3 license.
 
@@
 This Nickel file is then \emph{verified} against the contract (\todo{link to contract}) and translated to a JSON file fed to \ecg.
-The contract is versionned to take into account the potential future changes to the descriptions and the contract.
+The contract is versioned to take into account the potential future changes to the descriptions and the contract.
 
@@
 Hence, we capture the result of the download at the given link (success or error).
-In case of a succesful download, we also compute the cryptographic hash of the result of download to capture any variation in the source of the artifact which could explain later variations in the resulting software environment.
+In case of a successful download, we also compute the cryptographic hash of the result of download to capture any variation in the source of the artifact which could explain later variations in the resulting software environment.
 
@@
 This allows us to trace the version of the package.
-In the case of a simple downloading of the sources of a package (\texttt{misc\_packages} in Listing 1), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being dowloaded.
+In the case of a simple downloading of the sources of a package (\texttt{misc\_packages} in Listing 1), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
 The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
-Hence, \ecg\ will download \emph{localy} the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
+Hence, \ecg\ will download \emph{locally} the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
 
@@
 
-In this section, we present the initial protocole of our longitudinal study.
+In this section, we present the initial protocol of our longitudinal study.
 
@@
 
-Four out the five articles provided a Zenodo archive to the artifact, the other provided a github link with a specified \texttt{git} tag.
+Four out the five articles provided a Zenodo archive to the artifact, the other provided a GitHub link with a specified \texttt{git} tag.
 
@@
     The color of the bar corresponds to the month when a specific version of a package has been introduced in the software environment.
-    We can see that the proportion of package versions similar to the versions in the initial build is decreasing throught the months.
+    We can see that the proportion of package versions similar to the versions in the initial build is decreasing through the months.
   }
@@
 
-While this initial study only considered five \df s, it releaved that \textbf{the software environments produced by the considered \df s variate \emph{every month}} which highlights the lack of reproducibility of such practices.
+While this initial study only considered five \df s, it revealed that \textbf{the software environments produced by the considered \df s variate \emph{every month}} which highlights the lack of reproducibility of such practices.
 The question of the significance of these results on such a small sample will need to be answered in a larger scale study.
@@
 However, this practice raises serious questions in terms of scalability and sustainability in terms of storage.
-The lack of reproducibility of \df s shoud not be hidden with long-term storage: the practices from the conferences, journals, and the authors need to change.
+The lack of reproducibility of \df s should not be hidden with long-term storage: the practices from the conferences, journals, and the authors need to change.
 

### 2025-03-24 src/main.tex
[text] software heritage and zenodo links

@@
 In this section, we present \ecg, a tool for the periodic building of \df s from research artifacts and the collection of software environment information.
-\ecg\ is a 280-line Python3 script under the GPLv3 license.
+\ecg\ is a 280-line Python3 script~\cite{ecg} under the GPLv3 license.
 
@@
 One of it's feature is the possibility to ``verify'' that an configuration is ``well-formed'' with a \emph{contract}.
-This Nickel file is then \emph{verified} against the contract (\todo{link to contract}) and translated to a JSON file fed to \ecg.
+This Nickel file is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file fed to \ecg.
 The contract is versioned to take into account the potential future changes to the descriptions and the contract.
@@
 %\todo{when was the artifact submission, and publicaiton of the papers for europar?}
+The data collected is available on Zenodo~\cite{dataset_europar24}.
 

### 2025-03-26 src/main.tex
[text] first review q

@@
 \begin{abstract}
-  The reproducibility crisis struck all the fields of Science, including Computer Science.
+  The reproducibility crisis struck all the fields of Science, including Computer Science (CS).
   To tackle the issue, the CS community set up artifact evaluation processes in conferences and journals to evaluate the reproducibility of the results shared in the publications.
-  Authors thus share their artifacts included code, data and software environment required to reproduce the results.
-  One method for sharing the software environment, that has been ``advertised'' by conferences and journals, is to use container technologies such a Docker and Apptainer.
-  However, these tools relies on non-reproducible tools, hence producing non-reproducible containers.
+  Authors thus share their artifacts to reviewers, including code, data and software environment required to reproduce the results.
+  One method for sharing the software environment that has been proposed by conferences and journals, is to use container technologies such a Docker and Apptainer.
+  However, these tools rely on non-reproducible tools, hence producing non-reproducible containers.
 
-  In this paper, we present a tool and methodology to evaluate the variation in software environments through time of container images from research artifacts.
+  In this paper, we present a tool and methodology to evaluate the variations through time in software environments of container images from research artifacts.
   We also present initial results on a small set of \df s from the EuroPar'24 conference.
@@
 Then, artifact reviewers would try to reproduce those main results given the artifact from the authors.
-Based on the result of the reproduction attempt, the article is then awarded \emph{badges} to promote its reproducibility aspect.
+Based on the result of the reproduction attempt, the article is then awarded \emph{badges} to promote its reproducibility aspect~\cite{acm-badges}.
 The link to the artifacts can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, workflow)~\cite{paperswithcode, kang2023papers, hong2013software}.
@@
 Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate}.
-To tackle this issue, the conference \ae\ committees have been pushing the authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment.
-However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
-Moreover, in the case of usual virtualization techniques to generate containers or virtual machines, users have to write the recipe made of a list of commands to execute from a base image.
+To tackle this issue, the conference \adae\ committees have been asking the authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment.
+%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
+In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the recipe made of a list of commands to execute from a base image.
 This approach has several limitations.
@@
 Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
-Hence, rebuilding a recipe of a container or a virtual machine can yield different results, and thus ``breaking'' the reproducibility of the results.
-One solution would be to store the container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
+Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus break the reproducibility of the results.
+One solution would be for the authors to store the built container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
 But this come with the downside of requiring more storage space than just storing the recipe (binary vs. text file).
-This practice has thus some \textbf{severe scalability and sustainability issues in terms of longevity}.
+This practice has thus some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{guilloteau2024longevity}.
 Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
@@
 
-An \df\ from an artifact is represented as a Nickel~\cite{nickel} file.
+With \ecg, a \df\ from an artifact is represented as a Nickel configuration file~\cite{nickel}.
 Nickel is a configuration language with functional paradigm.
-One of it's feature is the possibility to ``verify'' that an configuration is ``well-formed'' with a \emph{contract}.
-This Nickel file is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file fed to \ecg.
+One of its features is the possibility to \emph{verify} that a configuration is valid with respect to a \emph{contract}.
+The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file fed to \ecg.
 The contract is versioned to take into account the potential future changes to the descriptions and the contract.
+Listing~\ref{lst:nickel} shows an example of Nickel representation.
 
-\begin{lstlisting}[caption=Example of Artifact Nickel representation]
+\begin{lstlisting}[caption=Example of Artifact Nickel representation\label{lst:nickel}]
 {
@@
 Hence, we capture the result of the download at the given link (success or error).
-In case of a successful download, we also compute the cryptographic hash of the result of download to capture any variation in the source of the artifact which could explain later variations in the resulting software environment.
+In case of a successful download, we also compute the cryptographic hash of the result of the download to capture any variation in the source of the artifact which could explain later variations in the resulting software environment.
 
@@
 In the case of a \df s failing to build, we capture the error (missing image, failing command, etc.).
-In fact, we are only able to capture the first error of the build even if there might be several errors in the building process.
+In fact, we are only able to capture the \emph{first error} of the build even if there might be several errors in the building process.
 The Docker error is extracting by \texttt{grep}ing the building process log.
@@
 \ecg\ is able to capture these different ways to install packages.
-Hence, for each package, we extract the package name, the version (the \texttt{git} commit if using \texttt{git}), and the tool used to install the package.
+Hence, for each package, we extract the package name, the version, and the tool used to install the package.
 
-As of for now, \ecg\ is able to query information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are install by source (\texttt{package\_manager} in Listing 1).
-In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing 1), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container.
+As of for now, \ecg\ is able to query information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are install by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}).
+In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing~\ref{lst:nickel}), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container.
 This allows us to trace the version of the package.
-In the case of a simple downloading of the sources of a package (\texttt{misc\_packages} in Listing 1), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
+In the case of a simple download of the sources of a package (\texttt{misc\_packages} in Listing~\ref{lst:nickel}), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
 The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
@@
 This framework leaves a lot of room for contributions on the writing of the Nickel descriptions files.
-The community can contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.
+The community could contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.
 
@@
 \item For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
-\item From the artifact and the \df, we write the associated Nickel file: where is stored the artifact? how to extract the \df\ out of the artifact? which package managers are used?
+\item From the artifact and the \df, we write the associated Nickel file (Listing~\ref{lst:nickel}): where is stored the artifact?
+how to extract the \df\ out of the artifact?
+which package managers are used?
 \item Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
@@
 \item Each container has an allocated time of 10 minutes to build.
-\item \ecg\ is executed on the (x86) machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
+\item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
 \end{itemize}
@@
 
-This section presents the information we would like to extract from the data gathered.
+%This section presents the information we would like to extract from the data gathered.
 
@@
 
-We wrote the description in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and built the associated \df s once every month for six months.
+We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build the associated \df s once every month for six months.
 % "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
@@
 We can see that most of the container used the \texttt{ubuntu} base image and that \emph{all} the \df s specified the version of the base image to use, thus limiting the variations in terms of produced software environment.
-However, all of the \df s called the \texttt{apt-get update} command, which fetches every time the latest versions of the packages and thus does introduces variations.
+However, all of the \df s called the \texttt{apt-get update} command, which fetches every time the latest versions of the packages and thus does introduce variations.
 
@@
 We presented in Section~\ref{sec:ecg} \ecg, and its workflow (Section~\ref{sec:workflow}), a tool to download, build and extract information from research artifacts \df.
-Section~\ref{sec:protocol} presented the protocol to answer our scientific questions.
-We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every months for six months.
+Section~\ref{sec:protocol} presented the protocol to observe the potential variations in software environment through time.
+We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every month for six months.
 
@@
 
+\textbf{
   A practice to deal with the lack of reproducibility of the \df s is for the authors to store the built container onto some long-time storage such as Zenodo.
@@
   The lack of reproducibility of \df s should not be hidden with long-term storage: the practices from the conferences, journals, and the authors need to change.
+}
 
@@
 
-This paper aimed to be an initial version of the design of a future larger study around the longevity of containers.
+This paper aims to be an initial version of the design of a future larger study around the longevity of containers.
 We will consider not only Docker containers, but also other containerization tools such as Apptainer.

### 2025-04-03 src/main.tex
co-authors
[text] improving

@@
   % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
-  Longitudinal Study of the Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
+  Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
 }
 
+% \author{Quentin Guilloteau}
+% \email{Quentin.Guilloteau@inria.fr}
+% \affiliation{%
+%   \institution{INRIA}
+%   \country{France}
+% }
+
 \author{Quentin Guilloteau}
+\orcid{0009-0003-7645-5044}
 \email{Quentin.Guilloteau@inria.fr}
 \affiliation{%
-  \institution{INRIA}
+  \institution{Inria, CNRS, ENSL, UCBL1, LIP}
+  \city{Lyon}
   \country{France}
+}
+
+\author{Antoine Waehren}
+\email{A.Waehren@stud.unibas.ch}
+\affiliation{%
+  \institution{University of Basel}
+  \city{Basel}
+  \country{Switzerland}
+}
+
+\author{Florina M. Ciorba}
+\email{Florina.Ciorba@unibas.ch}
+\affiliation{%
+  \institution{University of Basel}
+  \city{Basel}
+  \country{Switzerland}
 }
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Reproducibility, Artifact Evaluation, Badges, Lifetime}
+\keywords{Artifact Evaluation, Containers, Longevity, Sustainability}
 
@@
 The entire scientific community has been struck by the ``Reproducibility Crisis''~\cite{baker500ScientistsLift2016}, and Computer Science does not make exception~\cite{collberg_repeatability_2015}.
-In order to improve the reproducibility of articles in Computer Science, the community set up \adae\ processes~\cite{kidwell2016badges} where authors would submit (if they want) the artifact required to reproduced the main results from their article.
+In order to improve the reproducibility of articles in Computer Science, the community set up \adae\ processes~\cite{kidwell2016badges} where authors would \emph{voluntary} submit the artifact required to reproduced the main results from their article.
 Then, artifact reviewers would try to reproduce those main results given the artifact from the authors.
 Based on the result of the reproduction attempt, the article is then awarded \emph{badges} to promote its reproducibility aspect~\cite{acm-badges}.
-The link to the artifacts can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, workflow)~\cite{paperswithcode, kang2023papers, hong2013software}.
+The link to the artifact can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, workflow)~\cite{paperswithcode, kang2023papers, hong2013software}.
 
 One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
-Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate}.
-To tackle this issue, the conference \adae\ committees have been asking the authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment.
+Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate} \todo{more}.
+To tackle this issue, the conference \adae\ committees have been advising authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
 %However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
-In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the recipe made of a list of commands to execute from a base image.
+In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' made of a list of commands to execute from a ``base image'' (for example, \df\ for Docker containers).
 This approach has several limitations.
@@
 Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
-Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus break the reproducibility of the results.
+Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus threatens the reproducibility of the results.
 One solution would be for the authors to store the built container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
-But this come with the downside of requiring more storage space than just storing the recipe (binary vs. text file).
-This practice has thus some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{guilloteau2024longevity}.
+But this come with the downside of requiring more storage space than just storing the recipe (binary vs.
+text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{guilloteau2024longevity}.
 Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
@@
 
-In this paper, we propose our \textbf{initial design of a longitudinal study of the variations through time of the software environment produced by \df s from research artifacts.}
-We show initial results from five \df s extracted from the EuroPar'24 conference reproducibility artifacts built every month for six months.
+In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
+We show initial results from five \df s extracted from the EuroPar'24 conference reproducibility artifacts built every month for seven months.
 % \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}
@@
 
-The following of this paper is organized as follows:
-Section~\ref{sec:ecg} presents our software framework for this longitudinal study.
+The remainder of this paper is organized as follows:
+Section~\ref{sec:ecg} presents \ecg, our software framework for this longitudinal study.
 The protocol is presented in Section~\ref{sec:protocol}.
@@
 
-In this section, we present \ecg, a tool for the periodic building of \df s from research artifacts and the collection of software environment information.
+In this section, we present \ecg, a tool for building \df s from research artifacts and collecting information on their produced software environment.
 \ecg\ is a 280-line Python3 script~\cite{ecg} under the GPLv3 license.
@@
 Nickel is a configuration language with functional paradigm.
-One of its features is the possibility to \emph{verify} that a configuration is valid with respect to a \emph{contract}.
-The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file fed to \ecg.
+One of Nickel's features is the possibility to \emph{verify} that a configuration file is valid with respect to a \emph{contract}.
+The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file given as input to \ecg.
 The contract is versioned to take into account the potential future changes to the descriptions and the contract.
@@
 
-\begin{lstlisting}[caption=Example of Artifact Nickel representation\label{lst:nickel}]
+\begin{lstlisting}[caption=Example of Artifact Nickel representation, label=lst:nickel]
 {
@@
 
+This section presents the different information collected by \ecg, and their motivations.
+
 \subsubsection{Hash of the Artifact}
@@
 Hence, we capture the result of the download at the given link (success or error).
-In case of a successful download, we also compute the cryptographic hash of the result of the download to capture any variation in the source of the artifact which could explain later variations in the resulting software environment.
+In case of a successful download, we also compute the cryptographic hash of the result of the download to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
 
@@
 In fact, we are only able to capture the \emph{first error} of the build even if there might be several errors in the building process.
-The Docker error is extracting by \texttt{grep}ing the building process log.
+The Docker error is extracted by \texttt{grep}ing the building process log.
 
@@
 
-The software environment can be composed of several packages.
-Theses packages can have been installed by different means.
+The software environment can be composed of several packages, which can have been installed by different means.
 The most common way is through package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
@@
   \centering
-  \fbox{
-  \includegraphics[width=\textwidth]{./figs/workflow.pdf}
-  }
+  %\fbox{
+  \includegraphics[width=1\textwidth,trim={1.40cm 1.65cm 1.65cm 1.65cm},clip]{./figs/workflow.pdf}
+  %}
   \caption{
@@
 The workflow of \ecg\ is depicted in Figure~\ref{fig:workflow}.
-We make sure to use an empty local Docker cache before building the containers.
+% We make sure to use an empty local Docker cache before building the containers.
 The workflow itself is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment is managed with Nix~\cite{dolstra_nix_2004}.
@@
 The community could contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.
+\todo{more}
 
@@
 \item Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
-\item We run \ecg\ the closer to the start of each month as possible.
+%\item We run \ecg\ the closer to the start of each month as possible.
 \item The local Docker cache of these machine is cleaned before the building of each container.
-\item We will execute \ecg\ on each artifact every month for a full year (13 months).
+\item We will execute \ecg\ on each artifact every month, as close as possible to the start of the month, for a full year (\ie\ 13 months).
 \end{enumerate}
@@
 \begin{itemize}
+\item We make sure to use an empty local Docker cache before building the containers.
 \item Each container has an allocated time of 10 minutes to build.
@@
 
-We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build the associated \df s once every month for six months.
+We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build the associated \df s once every month for seven months.
 % "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
-The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, and March 12$^{th}$ 2025.
+The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
 The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
 %\todo{when was the artifact submission, and publicaiton of the papers for europar?}
-The data collected is available on Zenodo~\cite{dataset_europar24}.
+The data collected is available on Zenodo~\cite{dataset_europar24}. \todo{update zenodo with the april measures}
 
@@
   \centering
-  \scalebox{0.925}{
+  \scalebox{0.850}{
   \begin{tabular}{llll}
@@
     \midrule
-    \texttt{canon\_solving} & \texttt{ubuntu} & \texttt{22.04} & Yes \\
-    \texttt{geijer\_how}    & \texttt{ubuntu} & \texttt{22.04} & Yes \\
-    \texttt{hiraga\_peanuts}& \texttt{devcontainers/cpp} & \texttt{1-debian-12} & Yes \\
-    \texttt{munoz\_fault}   & \texttt{ubuntu} & \texttt{22.04} & Yes \\
-    \texttt{wolff\_fast}    & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{canon\_solving}~\cite{canon_solving}   & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{geijer\_how}~\cite{geijer_how}         & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{hiraga\_peanuts}~\cite{hiraga_peanuts} & \texttt{devcontainers/cpp} & \texttt{1-debian-12} & Yes \\
+    \texttt{munoz\_fault}~\cite{munoz_fault}       & \texttt{ubuntu} & \texttt{22.04} & Yes \\
+    \texttt{wolff\_fast}~\cite{wolff_fast}         & \texttt{ubuntu} & \texttt{22.04} & Yes \\
     \bottomrule
@@
   }
-  \caption{Information about the \df s from the study}\label{tab:info}
+  \caption{Information about the \df s from the study.}\label{tab:info}
 \end{table}
@@
 We can see that even after \emph{a single month} of the initial build, the software environment is already different for \emph{all} of the studied artifacts.
-This is a severe threat to reproducibility as one month is the time-frame for a artifact evaluation process in conferences.
+This is a severe threat to reproducibility as one month is the time-frame for an artifact evaluation process in conferences.
 This means that the \df\ submitted by the authors can generate different software environment between the submission of the artifact and the reproduction attempt of the reviewers.
@@
 
-Figure~\ref{fig:results_per_tool} depicts the evolution of the package version managed by each tool: \texttt{dpkg}, \texttt{git}, \texttt{pip} and manual download and build (\texttt{misc}).
+Figure~\ref{fig:results_per_tool} depicts the evolution of the package versions managed by each tool present in the studied artifact: \texttt{dpkg}, \texttt{git}, \texttt{pip}, and manual download and build (\texttt{misc}).
 We can see that the tools introducing variation in the software environment are \texttt{dpkg} and \texttt{pip}.
 
+\todo{update with april measures}
 The packages that changed the most in terms of versions are:
@@
 This paper presented the initial design of our longitudinal study of the evolution of the software environments produced by \df s.
-We presented in Section~\ref{sec:ecg} \ecg, and its workflow (Section~\ref{sec:workflow}), a tool to download, build and extract information from research artifacts \df.
+We presented in Section~\ref{sec:ecg} \ecg, and its workflow (Section~\ref{sec:workflow}), a tool to download, build and extract information from research artifacts \df s.
 Section~\ref{sec:protocol} presented the protocol to observe the potential variations in software environment through time.
-We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every month for six months.
+We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every month for seven months.
 

### 2025-04-03 scripts/results_per_artifact.R

@@ -26,13 +26,14 @@ p <- df %>%
 	ggplot(aes(x = factor(date), y = 100 * total / total_pkgs, fill = rank)) +
 	geom_col() +
 	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 5, label = total), angle=0, size=2.0) +
-	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 10, label = paste("(",round(100*total/total_pkgs, digits = 1), "%)",sep="")), angle=0, size=1.5) +
+	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 15, label = paste("(",round(100*total/total_pkgs, digits = 1), "%)",sep="")), angle=0, size=1.25) +
 	ylab("Packages in environment [%]") +
 	facet_wrap(~artifact, ncol = 5) +
-	ggtitle("Evolution of the packages versions through time") +
-	scale_fill_grey("Month when the package version was introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
+	ggtitle("Evolution of the packages versions through time for each Dockerfile studied") +
+	scale_fill_grey("Month when the package version\nwas introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
 	scale_x_discrete("Months after initial build", labels=append(c("Initial"), seq(length(dates)-1))) +
 	theme_bw() +
-	theme(legend.position = "bottom", strip.background = element_blank())
+	theme(legend.position = "bottom", strip.background = element_blank()) +
+	guides(fill = guide_legend(reverse=TRUE, nrow=1))
 
-ggsave(plot=p,outfile, width=8, height=3.25)
+ggsave(plot=p,outfile, width=8, height=3)

### 2025-04-03 scripts/results_per_tool.R

@@ -30,12 +30,15 @@ p <- df %>%
 	ylab("Packages in environment [%]") +
 	facet_wrap(~tool, ncol = 2) +
 	ggtitle("Evolution of the packages versions through time") +
-	scale_fill_grey("Month when the package version\nwas introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
+	scale_fill_grey("Month when the package version was introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
 	scale_x_discrete("Months after initial build", labels=append(c("Initial"), seq(length(dates)-1))) +
+	guides(fill = guide_legend(reverse=TRUE, nrow=1, title.position ="top", theme = theme(
+	  legend.title = element_text(size = 9.2)
+	  ))) +
 	theme_bw() +
         theme(legend.position = "bottom", strip.background = element_blank(),
           plot.title.position = "plot",
           plot.caption.position =  "plot",
           legend.location = "plot")
 
-ggsave(plot=p,outfile, width=4, height=4.25)
+ggsave(plot=p,outfile, width=4, height=3.75)

### 2025-04-04 src/main.tex
[text] table 1

@@
 One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
-Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate} \todo{more}.
+Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate, sokolowski2024impact}.
 To tackle this issue, the conference \adae\ committees have been advising authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
@@
 But this come with the downside of requiring more storage space than just storing the recipe (binary vs.
-text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{guilloteau2024longevity}.
+text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
 Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
@@
 % We make sure to use an empty local Docker cache before building the containers.
-The workflow itself is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment is managed with Nix~\cite{dolstra_nix_2004}.
+The workflow itself \cite{ecg_workflow} is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment \cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.
 
@@
 %\todo{when was the artifact submission, and publicaiton of the papers for europar?}
-The data collected is available on Zenodo~\cite{dataset_europar24}. \todo{update zenodo with the april measures}
+The data collected is available on Zenodo~\cite{dataset_europar24}.
 
@@
   \centering
-  \scalebox{0.850}{
+  \scalebox{0.800}{
   \begin{tabular}{llll}
     \toprule
-    Artifact & Base image & Version & \texttt{update}?\\
+	\multirow{2}{*}{Artifact} & \multicolumn{2}{c}{Docker Base Image used} & \multirow{2}{*}{
+					\begin{tabular}{c}Calling\\ \texttt{apt update}? \end{tabular}
+}\\
+\cmidrule(lr){2-3}
+     & Name & Version & \\
     \midrule
@@
 
-\todo{update with april measures}
 The packages that changed the most in terms of versions are:
 \begin{enumerate}
-\item \texttt{linux-libc-dev:amd64} (installed via \texttt{dpkg}) with 6 different versions for each artifact (1 per new build attempt)
+\item \texttt{linux-libc-dev:amd64} (installed via \texttt{dpkg}) with 7 different versions for each artifact (1 per new build attempt), except for \texttt{hiraga\_peanuts} where there were 4 different versions.
+\item \texttt{numpy} (installed via \texttt{pip}) with 6 different versions for the artifact \texttt{wolff\_fast}
 \item \texttt{fonttools} (installed via \texttt{pip}) with 5 different versions for the artifacts \texttt{geijer\_how} and \texttt{wolff\_fast}
-\item \texttt{numpy} (installed via \texttt{pip}) with 5 different versions for the artifact \texttt{wolff\_fast}
 \end{enumerate}

### 2025-04-04 src/main.tex
[text] submittable version

@@
 The entire scientific community has been struck by the ``Reproducibility Crisis''~\cite{baker500ScientistsLift2016}, and Computer Science does not make exception~\cite{collberg_repeatability_2015}.
-In order to improve the reproducibility of articles in Computer Science, the community set up \adae\ processes~\cite{kidwell2016badges} where authors would \emph{voluntary} submit the artifact required to reproduced the main results from their article.
+In order to improve the reproducibility of articles in Computer Science, the community set up, in conferences and journals, \adae\ processes~\cite{kidwell2016badges} where authors would \emph{voluntary} submit the artifact required to reproduced the main results from their article.
 Then, artifact reviewers would try to reproduce those main results given the artifact from the authors.
@@
 One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
-Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate, sokolowski2024impact}.
+Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
 To tackle this issue, the conference \adae\ committees have been advising authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
 %However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
-In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' made of a list of commands to execute from a ``base image'' (for example, \df\ for Docker containers).
+In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
 This approach has several limitations.
@@
 Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus threatens the reproducibility of the results.
-One solution would be for the authors to store the built container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
+One solution would be for the authors to store, and then provide, the built container itself (\ie\ the tarball) onto some long term storage (for example Zenodo~\cite{zenodo}).
 But this come with the downside of requiring more storage space than just storing the recipe (binary vs.
@@
 The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file given as input to \ecg.
-The contract is versioned to take into account the potential future changes to the descriptions and the contract.
 Listing~\ref{lst:nickel} shows an example of Nickel representation.
+The contract is versioned (line 2 of Listing~\ref{lst:nickel}) to take into account the potential future changes to the descriptions and the contract.
 
@@
 
-\subsubsection{Hash of the Artifact}
+\subsubsection{Cryptographic Hash of the Artifact}
 
-The articles give the link to the artifact.
-This artifact can be shared in several ways, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
-Hence, we capture the result of the download at the given link (success or error).
-In case of a successful download, we also compute the cryptographic hash of the result of the download to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
+Artifacts can be shared in several ways by the authors, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
+Hence, \ecg\ captures the result of the download at the given link (success or error).
+In case of a successful download, \ecg\ computes the cryptographic hash of the result of the download to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
 
@@
 
-In the case of a \df s failing to build, we capture the error (missing image, failing command, etc.).
-In fact, we are only able to capture the \emph{first error} of the build even if there might be several errors in the building process.
+In the case of a \df\ failing to build, \ecg\ capture the error returned by the building process (\eg\ missing image, failing command).
+In fact, \ecg\ is only able to capture the \emph{first error} of the build even if there might be several errors in the building process.
 The Docker error is extracted by \texttt{grep}ing the building process log.
@@
 \ecg\ is able to capture these different ways to install packages.
-Hence, for each package, we extract the package name, the version, and the tool used to install the package.
+Hence, for each package, \ecg\ extracts the package name, the version, and the tool used to install the package.
 
-As of for now, \ecg\ is able to query information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are install by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}).
-In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing~\ref{lst:nickel}), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container.
-This allows us to trace the version of the package.
-In the case of a simple download of the sources of a package (\texttt{misc\_packages} in Listing~\ref{lst:nickel}), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
+As of for now, \ecg\ is able to query information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are install by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}), or the case of packages installed in virtual Python environments.
+In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing~\ref{lst:nickel}), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container, and use this commit as a ``version''.
+In the case of a download of the sources of a package (\texttt{misc\_packages} in Listing~\ref{lst:nickel}), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
 The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
-Hence, \ecg\ will download \emph{locally} the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
+Hence, \ecg\ will download \emph{locally} (outside of the container) the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
 
@@
   \centering
-  %\fbox{
-  \includegraphics[width=1\textwidth,trim={1.40cm 1.65cm 1.65cm 1.65cm},clip]{./figs/workflow.pdf}
-  %}
+  \includegraphics[width=1\textwidth,trim={1.37cm 1.65cm 1.65cm 1.65cm},clip]{./figs/workflow.pdf}
   \caption{
@@
     This JSON representation is then read by \ecg\ to
-    \textit{(i)} download the artifact,
-    \textit{(ii)} compute the hash of its content,
-    \textit{(iii)} build the container from the \df, and
-    \textit{(iv)} extract the software environment information from the built container.
+    download the artifact,
+    compute the hash of its content,
+    build the container from the \df, and
+    extract the software environment information from the built container.
     \ecg\ outputs files containing the information about the artifact and its \df\ .
@@
 The workflow of \ecg\ is depicted in Figure~\ref{fig:workflow}.
-% We make sure to use an empty local Docker cache before building the containers.
-The workflow itself \cite{ecg_workflow} is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment \cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.
+The workflow itself~\cite{ecg_workflow} is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment~\cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.
 
-\subsection{Contributions}
-
-This framework leaves a lot of room for contributions on the writing of the Nickel descriptions files.
-The community could contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.
-\todo{more}
+% \subsection{Contributions}
+% 
+% This framework leaves a lot of room for contributions on the writing of the Nickel descriptions files.
+% The community could contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.
+% \todo{more}
 
@@
 \item Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
-%\item We run \ecg\ the closer to the start of each month as possible.
 \item The local Docker cache of these machine is cleaned before the building of each container.
@@
 
-% \todo{}
-% 
-% \begin{itemize}
-% \item we convert the dockerfiles to nickel description
-% \item periodic rebuild of the dockerfiles
-% \item g5k
-% \item empty cache
-% \item max 10 mins
-% \item contributions
-% \end{itemize}
-
-% \newpage
 \section{Preliminary Results}\label{sec:results}
 
-%This section presents the information we would like to extract from the data gathered.
-
-% And there is no ``warning message'' that the environment is different.
-
-We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build the associated \df s once every month for seven months.
-% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
+We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build their associated \df s once every month for seven months.
+% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12" "2025-04-01"
 The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
-The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
-%\todo{when was the artifact submission, and publicaiton of the papers for europar?}
-The data collected is available on Zenodo~\cite{dataset_europar24}.
+The deadline for the authors to submit their artifacts at EuroPar'24 was on May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
+Hence our initial build took place almost 5 months after the artifacts submission.
 
 The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).
+The data collected, and presented in this section, is available on Zenodo~\cite{dataset_europar24}.
 
@@
 
-While this initial study only considered five \df s, it revealed that \textbf{the software environments produced by the considered \df s variate \emph{every month}} which highlights the lack of reproducibility of such practices.
-The question of the significance of these results on such a small sample will need to be answered in a larger scale study.
+While this initial study only considered five \df s, it revealed that \textbf{the software environments produced by the considered \df s variate \emph{every month}} which highlights the lack of reproducibility of such tools and methods to share software environments.
+The question of the results' significance on such a small sample will need to be answered in a larger scale study.
 
@@
   A practice to deal with the lack of reproducibility of the \df s is for the authors to store the built container onto some long-time storage such as Zenodo.
-  However, this practice raises serious questions in terms of scalability and sustainability in terms of storage.
-  The lack of reproducibility of \df s should not be hidden with long-term storage: the practices from the conferences, journals, and the authors need to change.
+  However, this practice raises serious questions in terms of scalability and sustainability in terms of storage~\cite{monroe2023preservation}.
+  The lack of reproducibility of \df s should not be hidden with long-term storage: the practices of the authors and of the \adae\ need to be re-evaluated.
 }
+
+The community might want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual reprenstation of their software environment which could be precisely reproduced, and with longevity guarentees~\cite{courtes2024source}.
+Moreover, Nix and Guix allow for the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.
 

### 2025-04-04 scripts/most_volatile_pkgs.R

@@ -1,0 +1,15 @@
+library(tidyverse)
+args = commandArgs(trailingOnly=TRUE)
+infile <- args[1]
+outfile <- args[2]
+
+df <- read_csv(infile, col_names = c("pkg", "version", "tool", "artifact", "date")) %>%
+	mutate(date = date(as_datetime(date)))
+
+df %>%
+    group_by(pkg, artifact, tool) %>%
+    summarize(different_versions = n_distinct(version)) %>%
+    ungroup() %>%
+    slice_max(different_versions, n=10) %>%
+    write_csv(outfile)
+

### 2025-04-04 src/main.tex
[text] fix typos

@@
 
-The community might want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual reprenstation of their software environment which could be precisely reproduced, and with longevity guarentees~\cite{courtes2024source}.
+The community might want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual representation of their software environment which could be precisely reproduced, and with longevity guarantees~\cite{courtes2024source}.
 Moreover, Nix and Guix allow for the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.

### 2025-04-07 src/main.tex
[text] florina's review on overleaf

--- 2025-04-04 src/main.tex
@@
 \usepackage{listings}
+
+\newcommand{\fmc}[1]{{\color{magenta}#1}} % Florina Ciorba
+
 \definecolor{commentcolour}{rgb}{0.04,0.43,0.17}
@@
 \begin{abstract}
-  The reproducibility crisis struck all the fields of Science, including Computer Science (CS).
-  To tackle the issue, the CS community set up artifact evaluation processes in conferences and journals to evaluate the reproducibility of the results shared in the publications.
-  Authors thus share their artifacts to reviewers, including code, data and software environment required to reproduce the results.
-  One method for sharing the software environment that has been proposed by conferences and journals, is to use container technologies such a Docker and Apptainer.
-  However, these tools rely on non-reproducible tools, hence producing non-reproducible containers.
+  The reproducibility crisis has affected all scientific disciplines, including computer science (CS).
+  To address this issue, the CS community has established artifact evaluation processes at conferences and in journals to evaluate the reproducibility of the results shared in publications.
+  Authors are therefore required to share their artifacts with reviewers, including code, data, and the software environment necessary to reproduce the results.
+  One method for sharing the software environment proposed by conferences and journals is to utilize container technologies such as Docker and Apptainer.
+  However, these tools rely on non-reproducible tools, resulting in non-reproducible containers.
 
-  In this paper, we present a tool and methodology to evaluate the variations through time in software environments of container images from research artifacts.
-  We also present initial results on a small set of \df s from the EuroPar'24 conference.
+  In this paper, we present a tool and methodology to evaluate variations over time in software environments of container images derived from research artifacts.
+  We also present initial results on a small set of \df s from the Euro-Par 2024 conference.
+
 \end{abstract}
@@
 
-The entire scientific community has been struck by the ``Reproducibility Crisis''~\cite{baker500ScientistsLift2016}, and Computer Science does not make exception~\cite{collberg_repeatability_2015}.
-In order to improve the reproducibility of articles in Computer Science, the community set up, in conferences and journals, \adae\ processes~\cite{kidwell2016badges} where authors would \emph{voluntary} submit the artifact required to reproduced the main results from their article.
+The scientific community has been hit by the ``Reproducibility Crisis''~\cite{baker500ScientistsLift2016}, and computer science makes no exception~\cite{collberg_repeatability_2015}.
+In order to improve the reproducibility of articles in Computer Science, the community set up, in conferences and journals, \adae\ processes~\cite{kidwell2016badges} where authors would \emph{voluntary} submit the artifact required to reproduce the main results from their article.
 Then, artifact reviewers would try to reproduce those main results given the artifact from the authors.
-Based on the result of the reproduction attempt, the article is then awarded \emph{badges} to promote its reproducibility aspect~\cite{acm-badges}.
+Based on the result of the reproduction attempt, the article is awarded \emph{badges} to promote its reproducibility aspect~\cite{acm-badges}.
 The link to the artifact can be found in the published version of the article, sometimes with an appendix giving more information about the artifact (\eg\ download instructions, hardware requirements, workflow)~\cite{paperswithcode, kang2023papers, hong2013software}.
@@
 Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
-To tackle this issue, the conference \adae\ committees have been advising authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
+To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
 %However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
@@
 First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
-These base images can also be removed by the administrators making all the recipes unbuildable~\cite{nvidia_cuda_lifetime}.
+Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
 Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
-Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus threatens the reproducibility of the results.
-One solution would be for the authors to store, and then provide, the built container itself (\ie\ the tarball) onto some long term storage (for example Zenodo~\cite{zenodo}).
-But this come with the downside of requiring more storage space than just storing the recipe (binary vs.
+Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus \fmc{endangers} the reproducibility of the results.
+One solution would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (for example, Zenodo~\cite{zenodo}).
+But this comes with the downside of requiring more storage space than just storing the recipe (binary vs.
 text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
@@
 In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
-We show initial results from five \df s extracted from the EuroPar'24 conference reproducibility artifacts built every month for seven months.
+We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months.
+\fmc{Could we give here the period October 1st, 2024 - April 1st, 2025 of these 7 months?}
 % \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}
@@
 
-The remainder of this paper is organized as follows:
+The remainder of this paper is organized as follows.
 Section~\ref{sec:ecg} presents \ecg, our software framework for this longitudinal study.
-The protocol is presented in Section~\ref{sec:protocol}.
-Section~\ref{sec:results} presents the initial results of a small scale version of the study.
+The study protocol is presented in Section~\ref{sec:protocol}.
+Section~\ref{sec:results} presents the initial results of a small-scale version of the study.
 Finally, we conclude and give perspectives in Section~\ref{sec:conclusion}.
@@
 
-In this section, we present \ecg, a tool for building \df s from research artifacts and collecting information on their produced software environment.
+In this section, we present \ecg, \fmc{Can we explain the acronym?} a tool for building \df s from research artifacts and collecting information on their resulting software environment.
 \ecg\ is a 280-line Python3 script~\cite{ecg} under the GPLv3 license.
@@
 With \ecg, a \df\ from an artifact is represented as a Nickel configuration file~\cite{nickel}.
-Nickel is a configuration language with functional paradigm.
+Nickel is a configuration language with a functional paradigm.
 One of Nickel's features is the possibility to \emph{verify} that a configuration file is valid with respect to a \emph{contract}.
 The Nickel file representing the artifact is then \emph{verified} against the contract~\cite{ecg_contract} and translated to a JSON file given as input to \ecg.
-Listing~\ref{lst:nickel} shows an example of Nickel representation.
-The contract is versioned (line 2 of Listing~\ref{lst:nickel}) to take into account the potential future changes to the descriptions and the contract.
+Listing~\ref{lst:nickel} shows an example of a Nickel representation.
+The contract is versioned (line 2 of Listing~\ref{lst:nickel}) to take into account potential future changes to the descriptions and the contract.
 
-\begin{lstlisting}[caption=Example of Artifact Nickel representation, label=lst:nickel]
+\begin{lstlisting}[caption=Example of Artifact representation in Nickel, label=lst:nickel]
 {
@@
 
-Artifacts can be shared in several ways by the authors, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
+The authors can share artifacts in several ways, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
 Hence, \ecg\ captures the result of the download at the given link (success or error).
-In case of a successful download, \ecg\ computes the cryptographic hash of the result of the download to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
+In case of a successful download, \ecg\ computes the cryptographic hash of the download result to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
 
-For instance, authors can give the link to a Zenodo archive that will \emph{always} point to the latest version of an artifact.
-Hence, the content of the artifact behind this link can change through time.
+For example, authors can link to a Zenodo archive that will \emph{always} point to the latest version of an artifact.
+Hence, the content of the artifact behind this link can change over time.
 
@@
 
-In the case of a \df\ failing to build, \ecg\ capture the error returned by the building process (\eg\ missing image, failing command).
-In fact, \ecg\ is only able to capture the \emph{first error} of the build even if there might be several errors in the building process.
+In the case of \df\ failing to build, \ecg\ captures the error returned by the building process (\eg\ missing image, failing command).
+In fact, \ecg\ is only capable of capturing the \emph{first error} of the build even if there might be several errors in the building process.
 The Docker error is extracted by \texttt{grep}ing the building process log.
@@
 The most common way is through package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
-But there are also installed packages for programming languages such as with \texttt{pip} for Python.
-Another way to install package is to download the sources and build them locally.
+But there are also packages installed for programming languages such as with \texttt{pip} for Python.
+Another way to install the package is to download the sources and build them locally.
 \ecg\ is able to capture these different ways to install packages.
@@
 
-As of for now, \ecg\ is able to query information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are install by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}), or the case of packages installed in virtual Python environments.
+Currently, \ecg\ is capable of querying information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are installed by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}), or the case of packages installed in virtual Python environments.
 In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing~\ref{lst:nickel}), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container, and use this commit as a ``version''.
@@
 The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
-Hence, \ecg\ will download \emph{locally} (outside of the container) the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
+Hence, \ecg\ will download \emph{locally} (outside the container) the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
 
@@
 
-The workflow of \ecg\ is depicted in Figure~\ref{fig:workflow}.
-The workflow itself~\cite{ecg_workflow} is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment~\cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.
+The workflow of \ecg\ is shown in Figure~\ref{fig:workflow}.
+The workflow itself~\cite{ecg_workflow} is governed by a Snakemake~\cite{koster2012snakemake} description, and its software environment~\cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.
 
@@
 \begin{enumerate}
-\item For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
-\item From the artifact and the \df, we write the associated Nickel file (Listing~\ref{lst:nickel}): where is stored the artifact?
-how to extract the \df\ out of the artifact?
-which package managers are used?
-\item Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
-\item The local Docker cache of these machine is cleaned before the building of each container.
-\item We will execute \ecg\ on each artifact every month, as close as possible to the start of the month, for a full year (\ie\ 13 months).
+\item For each conference with an artifact evaluation process, we examine all artifacts and consider only those that share their software environment using a \df.
+\item From the artifact and the \df we generate the associated Nickel file (Listing~\ref{lst:nickel}) which contains information such as the artifact's location, how to extract the \df\ from the artifact and the package managers used.
+%where is the artifact stored? how to extract the \df\ out from the artifact? which package managers are used?
+\item Once all Nickel descriptions for the artifacts of a conference have been completed, we use \ecg\ to construct the download, build, and extract information from each of them.
+\item Prior to building each container, the local Docker cache is cleared on these machines.
+\item We will execute \ecg\ on each artifact every month, as close to the beginning of the month as possible, for a full year (\ie\ 13 months).
 \end{enumerate}
@@
 \begin{itemize}
-\item We make sure to use an empty local Docker cache before building the containers.
-\item Each container has an allocated time of 10 minutes to build.
+\item We ensure that an empty local Docker cache is used before building the containers.
+\item Each container is allocated a 10-minute time frame to build.
 \item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
@@
 
-We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build their associated \df s once every month for seven months.
+We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the Euro-Par 2024 conference that used a Docker container to share their software environments, and used \ecg\ to build their associated \df s once every month for seven months.
 % "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12" "2025-04-01"
 The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
-The deadline for the authors to submit their artifacts at EuroPar'24 was on May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
-Hence our initial build took place almost 5 months after the artifacts submission.
+The deadline for the authors to submit their artifacts to Euro-Par 2024 was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
+Hence, our initial build took place almost 5 months after the artifacts submission.
 
 The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).
-The data collected, and presented in this section, is available on Zenodo~\cite{dataset_europar24}.
+The data collected and presented in this section is available on Zenodo~\cite{dataset_europar24}.
 
@@
 
-Four out the five articles provided a Zenodo archive to the artifact, the other provided a GitHub link with a specified \texttt{git} tag.
+Four out of the five articles provided a Zenodo archive to the artifact, the other provided a GitHub link with a specified \texttt{git} tag.
 
-No variation of the artifact sources, via the computation of their cryptographic hash, was observed.
+No variation of the artifact sources was observed, via computation of their cryptographic hash.
 
@@
 
-All the artifacts managed to successfully produce a container.
+All of the artifacts managed to successfully produce a container.
 
@@
 Table~\ref{tab:info} summarizes the information about the studied \df s.
-We can see that most of the container used the \texttt{ubuntu} base image and that \emph{all} the \df s specified the version of the base image to use, thus limiting the variations in terms of produced software environment.
-However, all of the \df s called the \texttt{apt-get update} command, which fetches every time the latest versions of the packages and thus does introduce variations.
+We can see that most of the container used the \texttt{ubuntu} base image and that \emph{all} the \df s specified the version of the base image to use, thus limiting the variations in terms of the software environment produced.
+However, all \df s called the \texttt{apt-get update} command, which fetches the latest versions of the packages every time and thus introduces variations.
 
@@
 
-Figure~\ref{fig:results_per_artifact} depicts the evolution through time of the software environment produced by the \df\ of each studied artifact.
-We can see that even after \emph{a single month} of the initial build, the software environment is already different for \emph{all} of the studied artifacts.
-This is a severe threat to reproducibility as one month is the time-frame for an artifact evaluation process in conferences.
-This means that the \df\ submitted by the authors can generate different software environment between the submission of the artifact and the reproduction attempt of the reviewers.
+Figure~\ref{fig:results_per_artifact} depicts the evolution through time of the software environment produced by \df\ of each artifact studied.
+We can see that even after \emph{a single month} of the initial build, the software environment is already different for \emph{all} of the artifacts studied.
+This is a severe threat to reproducibility as one month is the time frame for an artifact evaluation process in conferences.
+This means that the \df\ submitted by the authors can generate a different software environment between the submission of the artifact and the reproduction attempt of the reviewers.
 
-We can see that every month, new versions of packages are being introduced in the software environments of the containers.
+We can see that every month new versions of the packages are being introduced in the software environments of the containers.
 
@@
 
-Figure~\ref{fig:results_per_tool} depicts the evolution of the package versions managed by each tool present in the studied artifact: \texttt{dpkg}, \texttt{git}, \texttt{pip}, and manual download and build (\texttt{misc}).
-We can see that the tools introducing variation in the software environment are \texttt{dpkg} and \texttt{pip}.
+Figure~\ref{fig:results_per_tool} shows the evolution of the packages' versions managed by each tool present in the studied artifact: \texttt{dpkg}, \texttt{git}, \texttt{pip}, and manual download and build (\texttt{misc}).
+We can see that the tools that introduce variation in the software environment are \texttt{dpkg} and \texttt{pip}.
 
@@
 % In this paper, we presented an initial version of \ecg, and its workflow, a tool to study the variations through time of containers from research artifacts.
-% We evaluated the evolution of the software environment produced by \df\ from five artifacts from the EuroPar'24 conference.
+% We evaluated the evolution of the software environment produced by \df\ from five artifacts from the Euro-Par 2024 conference.
 % We observe that, even at this small scale, after only one month, there were variations in the produced software environments.
@@
 This paper presented the initial design of our longitudinal study of the evolution of the software environments produced by \df s.
-We presented in Section~\ref{sec:ecg} \ecg, and its workflow (Section~\ref{sec:workflow}), a tool to download, build and extract information from research artifacts \df s.
-Section~\ref{sec:protocol} presented the protocol to observe the potential variations in software environment through time.
-We showed initial results in Section~\ref{sec:results} on five artifacts from the EuroPar'24 conference built every month for seven months.
+We presented in Section~\ref{sec:ecg} \ecg, and its workflow (Section~\ref{sec:workflow}), a tool to download, build, and extract information from research artifacts \df s.
+Section~\ref{sec:protocol} presented the protocol to observe the potential variations in the software environment over time.
+We showed initial results in Section~\ref{sec:results} on five artifacts from the Euro-Par 2024 conference built every month for seven months.
 
-While this initial study only considered five \df s, it revealed that \textbf{the software environments produced by the considered \df s variate \emph{every month}} which highlights the lack of reproducibility of such tools and methods to share software environments.
-The question of the results' significance on such a small sample will need to be answered in a larger scale study.
+While this initial study only considered five \df s, it revealed that \textbf{the software environments produced by the considered \df s vary \emph{every month}} which highlights the lack of reproducibility of such tools and methods to share software environments.
+The question of the results' significance on such a small sample will need to be answered in a larger-scale study.
 
 \textbf{
-  A practice to deal with the lack of reproducibility of the \df s is for the authors to store the built container onto some long-time storage such as Zenodo.
+  A practice to deal with the lack of reproducibility of \df s is that the authors store the built container on long-term storage such as Zenodo.
   However, this practice raises serious questions in terms of scalability and sustainability in terms of storage~\cite{monroe2023preservation}.
-  The lack of reproducibility of \df s should not be hidden with long-term storage: the practices of the authors and of the \adae\ need to be re-evaluated.
+  The lack of reproducibility of \df s should not be hidden behind long-term storage: the practices of the authors and of \adae\ need to be re-evaluated.
 }
 
-The community might want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual reprenstation of their software environment which could be precisely reproduced, and with longevity guarentees~\cite{courtes2024source}.
-Moreover, Nix and Guix allow for the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.
+The community may want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual representation of their software environment which could be precisely reproduced and with longevity guarantees~\cite{courtes2024source}.
+Moreover, Nix and Guix allow the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.
 
-\paragraph{Perspectives}
+\paragraph{Future Perspectives}
 
-This paper aims to be an initial version of the design of a future larger study around the longevity of containers.
+%This paper aims to be an initial version of the design of a future larger study on the longevity of containers.
+% We will consider not only Docker containers, but also other containerization tools such as Apptainer.
+% We will thus write the descriptions of more artifacts from more conferences, aiming for around a hundreds total containers, from conferences of various fields of Science with various Computer Science technical levels.
+This paper serves as the foundation for the design of a larger future study on the longevity of containers. 
 We will consider not only Docker containers, but also other containerization tools such as Apptainer.
-We will thus write the descriptions of more artifacts from more conferences, aiming for around a hundreds total containers, from conferences of various fields of Science with various Computer Science technical levels.
+Therefore, we will write descriptions of more artifacts from more conferences, aiming at a total of approximately a hundred containers, from conferences spanning various scientific disciplines and with varying levels of computer science technical expertise.
 
@@
 
-This research was funded, in whole or in part, by the European Union’s Horizon 2020 research and innovation programme under grant agreement No. 957407 as DAPHNE.
+This research was funded in part by the European Union’s Horizon 2020 research and innovation programme under grant agreement No. 957407 (DAPHNE).
 

### 2025-04-07 src/main.tex
[text] answer florina's notes, and compression to fit on 4 pages

@@
 
-\newcommand{\fmc}[1]{{\color{magenta}#1}} % Florina Ciorba
+% \newcommand{\fmc}[1]{{\color{magenta}#1}} % Florina Ciorba
+\newcommand{\fmc}[1]{{}} % Florina Ciorba
 
@@
 Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
-Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus \fmc{endangers} the reproducibility of the results.
-One solution would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (for example, Zenodo~\cite{zenodo}).
+Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus endangers the reproducibility of the results.
+One would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (\eg\ Zenodo~\cite{zenodo}).
 But this comes with the downside of requiring more storage space than just storing the recipe (binary vs.
@@
 In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
-We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months.
-\fmc{Could we give here the period October 1st, 2024 - April 1st, 2025 of these 7 months?}
+We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months, from October 1$^{st}$ 2024 to April 1$^{st}$ 2025.
 % \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}
@@
 
-This section presents the different information collected by \ecg, and their motivations.
+This section presents the different information collected by \ecg.%, and their motivations.
 
@@
 
-In this section, we present the initial protocol of our longitudinal study.
+% In this section, we present the initial protocol of our longitudinal study.
+This section presents the initial protocol of our longitudinal study.
 

### 2025-04-07 src/main.tex
[text] double blind-ing
[text] compression

@@
   % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
-  Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
+  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
 }
@@
 
-\author{Quentin Guilloteau}
-\orcid{0009-0003-7645-5044}
-\email{Quentin.Guilloteau@inria.fr}
+% \author{Quentin Guilloteau}
+% \orcid{0009-0003-7645-5044}
+% \email{Quentin.Guilloteau@inria.fr}
+% \affiliation{%
+%   \institution{Inria, CNRS, ENSL, UCBL1, LIP}
+%   \city{Lyon}
+%   \country{France}
+% }
+% 
+% \author{Antoine Waehren}
+% \email{A.Waehren@stud.unibas.ch}
+% \affiliation{%
+%   \institution{University of Basel}
+%   \city{Basel}
+%   \country{Switzerland}
+% }
+% 
+% \author{Florina M. Ciorba}
+% \email{Florina.Ciorba@unibas.ch}
+% \affiliation{%
+%   \institution{University of Basel}
+%   \city{Basel}
+%   \country{Switzerland}
+% }
+
+\author{Anonymous Authors}
+\email{Anonymous email}
 \affiliation{%
-  \institution{Inria, CNRS, ENSL, UCBL1, LIP}
-  \city{Lyon}
-  \country{France}
-}
-
-\author{Antoine Waehren}
-\email{A.Waehren@stud.unibas.ch}
-\affiliation{%
-  \institution{University of Basel}
-  \city{Basel}
-  \country{Switzerland}
-}
-
-\author{Florina M. Ciorba}
-\email{Florina.Ciorba@unibas.ch}
-\affiliation{%
-  \institution{University of Basel}
-  \city{Basel}
-  \country{Switzerland}
+  \institution{Anonymous Institution}
+  \city{Anonymous City}
+  \country{Anonymous Country}
 }
@@
 To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
+In practice, about 12\% of the artifacts use such tools~\cite{guilloteau2024longevity}.
 %However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
@@
 We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months, from October 1$^{st}$ 2024 to April 1$^{st}$ 2025.
+We focus only on Docker containers as they are the most used in artifacts~\cite{guilloteau2024longevity}.
 % \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}
@@
 
-For example, authors can link to a Zenodo archive that will \emph{always} point to the latest version of an artifact.
+For example, authors can link to a Zenodo archive that will \emph{always} point to the latest artifact version.% of an artifact.
 Hence, the content of the artifact behind this link can change over time.
@@
 \paragraph{Details}
-
 \begin{itemize}
@@
 Hence, our initial build took place almost 5 months after the artifacts submission.
-
 The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).
@@
   \caption{
-    Evolution of the packages in the software environment of each container through time.
-    Each container has been rebuilt once a month.
+    Evolution of the packages in the software environment of each container over time.
+    %Each container has been rebuilt once a month.
     The color of the bar corresponds to the month when a specific version of a package has been introduced in the software environment.
-    We can see that the proportion of package versions similar to the versions in the initial build is decreasing through the months.
+    We can see that the decrease of the proportion of package versions similar to the versions in the initial build over time.
   }

### 2025-04-07 scripts/results_per_artifact.R

@@ -29,11 +29,11 @@ p <- df %>%
 	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 15, label = paste("(",round(100*total/total_pkgs, digits = 1), "%)",sep="")), angle=0, size=1.25) +
 	ylab("Packages in environment [%]") +
 	facet_wrap(~artifact, ncol = 5) +
-	ggtitle("Evolution of the packages versions through time for each Dockerfile studied") +
+	ggtitle("Evolution of the packages versions over time for each Dockerfile studied") +
 	scale_fill_grey("Month when the package version\nwas introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
 	scale_x_discrete("Months after initial build", labels=append(c("Initial"), seq(length(dates)-1))) +
 	theme_bw() +
 	theme(legend.position = "bottom", strip.background = element_blank()) +
 	guides(fill = guide_legend(reverse=TRUE, nrow=1))
 
-ggsave(plot=p,outfile, width=8, height=3)
+ggsave(plot=p,outfile, width=8, height=2.95)

### 2025-04-07 scripts/results_per_tool.R

@@ -29,7 +29,7 @@ p <- df %>%
 	geom_text(data = . %>% filter(as.character(rank) == dates[1]), aes(y = (100 * total/total_pkgs) - 15, label = paste("(",round(100*total/total_pkgs, digits = 1), "%)",sep="")), angle=0, size=1.5) +
 	ylab("Packages in environment [%]") +
 	facet_wrap(~tool, ncol = 2) +
-	ggtitle("Evolution of the packages versions through time") +
+	ggtitle("Evolution of the packages versions over time") +
 	scale_fill_grey("Month when the package version was introduced in the environment", labels=rev(append(c("Initial"), seq(length(dates)-1))))+
 	scale_x_discrete("Months after initial build", labels=append(c("Initial"), seq(length(dates)-1))) +
 	guides(fill = guide_legend(reverse=TRUE, nrow=1, title.position ="top", theme = theme(

### 2025-05-19 src/rebuttal.tex
start rebuttal

\documentclass[%
	11pt,
	final,
]{article}

% === PACKAGES IMPORT / CONFIGURATION =========================================

% --- ENCODING / FONTS -----------------

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{lmodern}  % vector fonts

% --- LAYOUT ---------------------------

\usepackage[%
	a4paper,
	margin=20mm,  % tighter margins
]{geometry}

% --- MATHEMATICS TYPESETTING ----------

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% \usepackage{amsthm}

% --- GRAPHICS / FIGURES ---------------

%\usepackage{cbfcolors}
\usepackage{graphicx}
\usepackage{standalone}
\usepackage{subcaption}
\usepackage{xcolor}

\usepackage{tikz}
% \usetikzlibrary{babel}  % avoid annoying interaction of babel with tikz
\usetikzlibrary{calc}
\usetikzlibrary{fit}

% --- LISTINGS / ALGORITHMS ------------

\usepackage{algorithm2e}  % pseudo-code

% --- ENHANCED TABLES ------------------

\usepackage{booktabs}  % publication quality tables

% --- CITATIONS ------------------------

\RequirePackage[%
	style=numeric,
	sortcites=true,
	giveninits=true,
	isbn=false,
]{biblatex}

\addbibresource{references.bib}

% --- MISC. PACKAGES -------------------

\usepackage[inline]{enumitem}  % enhanced enumerations
\usepackage[strict=true]{csquotes}  % advanced quotes
\usepackage{siunitx}  % correct units for physical quantities

\usepackage{mdframed}
\mdfdefinestyle{alert}{%
	frametitlealignment=\noindent,%
	linewidth=3pt,%
	hidealllines=true,%
	leftline=true,%
	rightline=true,%
}
\newmdenv[%
	style=alert,%
	linecolor=cyan,%
	backgroundcolor=cyan!10,%
	frametitle=Answer,%
]{review-answer}

% --- PDF SUPPORT ----------------------

\usepackage{hyperref}  % import as last package
\usepackage[capitalise]{cleveref}  % import after hyperref

% === MACROS DEFINITION =======================================================

% --- COMMON ABBREVIATIONS MACROS ------

% latin abbreviations, see:
%   - http://www.sussex.ac.uk/informatics/punctuation/capsandabbr/abbr
% comment by Sascha Hunold, see also:
%   - https://www.ieee.org/documents/style_manual.pdf
%     (p. 32, Short Reference List of Italics)
%   - http://web.ece.ucdavis.edu/~jowens/commonerrors.html

\newcommand{\eg}{e.g.\@}
\newcommand{\ie}{i.e.\@}
\newcommand{\cf}{cf.\@}
\newcommand{\circa}{ca.\@}
\newcommand{\perse}{\emph{per~se}}
\newcommand{\adhoc}{\emph{ad~hoc}}
\newcommand{\sic}{\emph{sic}}
\newcommand{\versus}{\emph{vs.\@}}
\newcommand{\aka}{a.k.a.\@}
\newcommand{\resp}{resp.\@}
\newcommand{\wrt}{w.r.t.\@}

\makeatletter
\newcommand{\etc}{etc\@ifnextchar.{}{.\@}}
\newcommand{\etal}{et~al\@ifnextchar.{}{.\@}}
\makeatother

\newcommand{\reviewitem}[2]{\paragraph{#1:} #2}

% === META DATA ===============================================================

\title{%
  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design
	\\
	{\Large%
	Answer to reviewers
	}
}

\author{Anonymous Authors}
\date{}

\begin{document}

\maketitle

This document is composed of three parts:

\begin{enumerate}
\item the detailed answers to the reviewers comments
\item a ``diff'' between the initial submission and the new submission (additions in underlined blue, deletion in crossed-out red)
\item the new submission
\end{enumerate}

\section{Review 1 (weak reject)}

Thank you for submitting your work at REP. Overall, focusing on reliability of reproducibility packages is a worthy endeavour, and well in-scope for this conference. The paper was also clearly written, however, I also have several concerns about this work, namely:

\begin{itemize}
\item The investigation is limited to only five artifacts from a single conference

\begin{review-answer}
  This initial study is indeed limited to five artifact from a single conference (Euro-Par24).
  The goal of this short paper is to investigate the necessity and potential of a larger scale study.
\end{review-answer}

\item Some aspects of the proposed system are unexplained and/or confusing.
  \begin{itemize}
   \item It is unclear what is the role of Nickel ad why there is need for an artifact representation in this framework - aren't Dockerfile sufficient to carry the same information?
\begin{review-answer}
  Nickel is simply a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''s.

  We cannot simply use/parse Dockerfiles to extract the same information.
  Indeed, the authors crafting their Dockerfile could do the installation of their software dependencies in a shell script that will then be called in the steps of the Dockerfile.
  The content of this script is unknown at the level of the Dockerfile.
  Hence the need to capture and describe all the downloads and installations to create the final Docker image.

  To improve clarity, we swapped Section 2.1 and Section 2.2, and argumented more the use of the Nickel representation.
\end{review-answer}
   \begin{itemize}
    \item Snakemake and Nix are mentioned without being introduced
      \begin{review-answer}
        We added a description of those tools in Section 2.3.
      \end{review-answer}
    \item From the description, the proposed toolchain appears capable to determine whether an environment can be built, but says nothing about whether experiments can actually be reproduced, which seems to miss the main problem with artifacts
      \begin{review-answer}
        That is correct, we only rebuild the Docker image, but do not reproduce the experiments.
        The reason is two-fold:
        \begin{itemize}
        \item The impact of software environment variations on experiments results have been heavily studied (see references [28, 32] ([27, 31] in the initial submission), cited in the second paragraph of the introduction)
        \item The energy-costs of actually reproducing the experiments for which the toolchain determined that the environment can be built must be weighed against the utility of reproducing those experiments. The goal of our short paper is to prove the concept of experiment building over time. This benefits researchers that indeed want/need to reproduce specific experiments and justify the associated energy-costs
        \end{itemize}

        We added a sentence in Section 3 to make this point more explicit.
      \end{review-answer}
    \end{itemize}
    \end{itemize}
\end{itemize}

Overall, while the idea is interesting, the approach and the evaluation thus have significant practical limitations.

\begin{review-answer}
  Indeed, the approach needs to be further developed and tested with a larger collection of artifacts, as we stated in the \emph{Perspectives} section.
  The extent of our current submission is a short-paper describing a proof of concept.
\end{review-answer}

\section{Review 2 (weak accept)}

\subsection{Overview}

This paper contributes to the litterature on Software Engineering, Reproducibility, and reproducibility of Sciences.

It introduces ecg, a tool used to rebuild Dockerfiles and analyse potential reproducibility issues.
It showcases the use of the tools by studying the evolution in time of 5 Dockerfile extracted from the Euro Par 2024 conference.

\subsection{Strenght}

\begin{itemize}
\item The topic is novel, relevant and in my opinion important. 
\item Contribution of a tool that can be used to conduct further experiments
\item Valuable first insights from the experiments
\end{itemize}

\subsection{Weeknesses}

\begin{itemize}
\item In my opinion the paper could be made more clear maybe by focusing on the most important elements: for example, you example that you use Nickel to reprensent artifacts, but it is unclear to me from the rest of the paper what is the added value for the experimental protocal. What happens if the Nickel ``contract'' is not respected ?
\begin{review-answer}
  Nickel is simply a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''.

  If the Nickel contract is not respected, then this means that the Nickel representation of the artifact's Dockerfile (Listing 1) does not respect the ``model'' (in the UML sense) / structure and content expected by \texttt{ecg} (the script performing the downloading, the building and the informtation extracting of the Docker image from the research artifact).
  The transformation from the Nickel representation to the JSON representation (given as input to \texttt{ecg}) will return an error, and the workflow will not even try to rebuild this Docker image.

  To improve clarity, we swapped Section 2.1 and Section 2.2, and argumented more the use of the Nickel representation.
\end{review-answer}
\end{itemize}

\subsection{Remarks}

\begin{itemize}
\item On the topic of software environment reproducibility, ``Reproducibility of Build Environments through Space and Time'' \url{https://dl.acm.org/doi/10.1145/3639476.3639767} would be good to cite.
  \begin{review-answer}
    We added a citation this paper in the Conclusion section (Reference [25]).
  \end{review-answer}
\end{itemize}

\section{Review 3 (accept)}

As a short paper, the authors evaluate five AD/AE appendices from EuroPar 2024 papers and study the Dockerfiles associated with these artifacts.
They study the variation of the software environment represented by the software and then report on the variation every month for a period of six months.
A key problem with the difference in software environments is that there is an apt-update command that refreshes the base image with updated packages.
It is not clear if this update - which is responsible for the variation reported - broke the software being studied, or if it still worked.
\begin{review-answer}
  This question has been studied in the litterature (see for example the citations [28, 32] ([27, 31] in the initial submission) cited in second paragraph of the introduction), and is not addressed in this paper.

We added a sentence in Section 3 to make this point more explicit.
\end{review-answer}
An easy fix would be to remove the apt-update and keeping the base image with a version.
\begin{review-answer}
  The preliminary results presented in this paper show that using tools such as \texttt{pip} also introduce variation in their produced software environment (see Figure 3).
\end{review-answer}
Other methods suggested include keeping a binary image as a Docker image or a Singularity image so we can freeze the environment for the paper.
\begin{review-answer}
  While this is true, we do not think that this approach is sustainable.
  Storing Gigabytes of binary container images on long-term storage is much more expensive than storing a few kilobytes text file.
  This point is discussed in the Introduction.
\end{review-answer}
There are some typos - mostly with spacing - that need to be fixed (latest artifact version.Hence -> version. Hence, Dockerfilewe -> Dockerfile we).
\begin{review-answer}
  The typos have been corrected.
\end{review-answer}
The tool developed by the authors (ecg) is provided with an anonymized URL and source code available for download with the GPLv3 license.

\section{Review 4 (weak reject)}

The paper "Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design" addresses an important and well-recognized issue in computational reproducibility: Docker containers, while widely used, are not inherently reliable for long-term archiving and reproducibility.

The authors confirm that independently building containers from the same Dockerfile can lead to different results due to changes in software versions and dependency drift, a problem previously documented in the reproducibility literature (e.g., "Reproducibility, Replicability, and Repeatability: A Survey of Reproducible Research with a Focus on High Performance Computing," \url{https://arxiv.org/pdf/2402.07530}).

\begin{review-answer}
The paper proposed by the reviewer does not document this issue: it claims (in the last paragraph of page 31) that there is an issue without showing any proof of it.
This is a common claim what we have seen in the litterature, but we have yet to see any proof of it.
Our work therefore aims to bring evidence supporting the claim that the reproducibility of Sofware Environments decays over time.

We added a sentence at the end of the introduction to explicit this gap in the litterature.
\end{review-answer}

In their study, the authors monitored Docker containers associated with published Euro-Par papers and observed that software dependencies inside containers do indeed evolve over time, potentially impacting the results produced by those containers.
As a solution, they propose an approach that combines using a textual description of the container build process and monitoring dependencies through tools such as Snakemake and Nix.

While I fully agree with the authors that Docker containers alone are insufficient to guarantee reproducibility, I have several concerns regarding the proposed solution.
Primarily, the suggested approach does not seem to fundamentally resolve the core issue of dependency management.
\begin{review-answer}
  There might be a misunderstanding.
  We do not propose another format to generate and share container images.
  We use the textual representation presented in Section 2.2 (Section 2.1 in the initial submission) to track what are the downloads and installations done in a Dockerfile from a research artifact.
  We then build the Docker image from the Dockerfile, and use our representation to perform the operation required to extract information about the generated software environment.
  We need this representation because we cannot simply use the Dockerfile to extract the information about which package is installed and how.
  There are some cases where authors write a shell scripts doing the installation of packages (\eg\ calls to \texttt{apt install}) and from the point of view of the Dockerfile these installation steps are not visible.
  Hence, we gather all the information about the downloads and installations in one textual representation: the one presented in Section 2.2 (Section 2.1 in the initial submission).

  To improve clarity, we swapped Section 2.1 and Section 2.2, and argumented more the use of the Nickel representation.
\end{review-answer}
Furthermore, the paper lacks a comparison with existing, widely adopted industrial practices for achieving reproducibility, particularly in domains like AI and machine learning.
\begin{review-answer}
  This is correct, and explicited in the \emph{Perspectives} section of the article.
  As this is a short paper and an initial design of what might be a larger scale experiment, we focused on a conference of our field (HPC), where we more used to the tools and practices.
  Previous study have shown that Docker is the main container tool used to produce container for research artifacts in HPC (see reference [20]), hence this initial focus on Docker.
\end{review-answer}

Current best practices in industry involve:

\begin{review-answer}
  There are indeed best practices than artifact authors can follow.
  Let us note that the practices of academia differ to the practices of industry, where the objectives are different.
  Having authors following better practices would be ideal, but unfortunately it is not the case.
\end{review-answer}

\begin{itemize}
\item Using stable base images and archiving them;
  \begin{review-answer}
    While this practice does improve long-term reproductiblity, it lacks scalability and sustainability as storing Gibabytes of binary is much more expensive than storing a few kilobytes of plain text.
    This point as been discussed in the conclusion section of the article.
  \end{review-answer}
\item Avoiding commands like "apt update" within Dockerfiles;
\item Explicitly specifying and fixing system-level dependencies using tools like apt, dpkg;
\item Managing Python dependencies precisely via pip, poetry, or uv;
  \begin{review-answer}
    From our understanding, this only guarentees the top-level dependencies to be correct, but if there are dependencies of those dependeices that have not been captured correctly, the final software environment might still differ.
  \end{review-answer}
\item Fetching code and artifacts using fixed hashes (e.g., via Git or direct downloads).
  \begin{review-answer}
    Using Git and direct downloads does not guarantees long-term reproducibility.
    The content behind a direct download can change while the URL stays the same.
    Git commits can be remove/rewritten and repositories deleted.

    This point has been discussed in the sections 2.1.1 and 2.1.3 (sections 2.2.1 and 2.2.3 in the initial submission).
  \end{review-answer}
\end{itemize}

These practices, when rigorously followed, can improve reproducibility within standard Docker workflows without introducing new tooling or formats.
\begin{review-answer}
  These practices might indeed improve reproducibility, while to garenteeing anything.
  We agree with the reviewer on the ``rigorously'' and ``improve'' terms.
  The issue is that it is nearly impossible to create a docker image that will be reproducible over time, but it is very easy to create a docker image that works now.
  So the reframed question is ``Are container recipes a real solution to the reproducibility problem like the calls for artifacts seem to claim?''.

  The bold text in the Conclusion section summarizes this point.
\end{review-answer}
Consequently, it appears that the reproducibility guarantees offered by the proposed approach could, in fact, be achieved using current methods without the added complexity of additional tools.
\begin{review-answer}
  We are not proposing a new format to build and share container images.
\end{review-answer}

Therefore, before introducing new tools and formats, it would strengthen the paper to include a detailed comparison with these existing reproducibility strategies and methodologies.
Such a comparison would help to clarify the unique advantages of the proposed method and better position it within the landscape of current solutions.

Overall, this paper identifies a significant problem and provides useful empirical data, but its proposed solution requires further development, particularly through a more rigorous evaluation against established practices.

\begin{review-answer}
  We fully acknowlege that a wider evaluation is required to target a full publication.
  As stated in the conclusion section (section 5), this short paper aims to be an initial version of a larger scale study.
\end{review-answer}

\end{document}

### 2025-05-19 src/for_reviewers.tex
start rebuttal

\documentclass{article}
\usepackage{pdfpages}
\usepackage{graphicx}
\begin{document}

\includepdf[pages=-]{rebuttal.pdf}
\section*{Changes made}
\includepdf[pages=-]{diff.pdf}
\section*{New version}
\includepdf[pages=-]{main.pdf}
\end{document}

### 2025-05-20 src/rebuttal.tex
aspell on the text

@@
 
-\title{%
-  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design
-	\\
-	{\Large%
-	Answer to reviewers
-	}
-}
-
-\author{Anonymous Authors}
-\date{}
+% \title{%
+%   SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design
+% 	\\
+% 	{\Large%
+% 	Answer to reviewers
+% 	}
+% }
+% 
+% \author{Anonymous Authors}
+% \date{}
 
@@
 
-\maketitle
-
-This document is composed of three parts:
-
-\begin{enumerate}
-\item the detailed answers to the reviewers comments
-\item a ``diff'' between the initial submission and the new submission (additions in underlined blue, deletion in crossed-out red)
-\item the new submission
-\end{enumerate}
+% \maketitle
 
@@
 \begin{review-answer}
-  Nickel is simply a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''s.
+  Nickel is a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''s.
 
   We cannot simply use/parse Dockerfiles to extract the same information.
-  Indeed, the authors crafting their Dockerfile could do the installation of their software dependencies in a shell script that will then be called in the steps of the Dockerfile.
+  Indeed, the authors crafting their Dockerfile could for example do the installation of their software dependencies in a shell script that will then be called in the steps of the Dockerfile.
   The content of this script is unknown at the level of the Dockerfile.
-  Hence the need to capture and describe all the downloads and installations to create the final Docker image.
+  Hence the need to capture manually the download and installations, and list them in this Nickel format.
 
-  To improve clarity, we swapped Section 2.1 and Section 2.2, and argumented more the use of the Nickel representation.
+  To improve clarity, we swapped Section 2.1 and Section 2.2, and explained more the use of the Nickel representation.
 \end{review-answer}
@@
 \begin{review-answer}
-  Indeed, the approach needs to be further developed and tested with a larger collection of artifacts, as we stated in the \emph{Perspectives} section.
+  Indeed, the approach needs to be further developed and tested with a larger collection of artifacts, as stated in the \emph{Perspectives} section.
   The extent of our current submission is a short-paper describing a proof of concept.
@@
 
-This paper contributes to the litterature on Software Engineering, Reproducibility, and reproducibility of Sciences.
+This paper contributes to the literature on Software Engineering, Reproducibility, and reproducibility of Sciences.
 
@@
 \begin{review-answer}
-  Nickel is simply a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''.
+  Nickel is a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''.
 
-  If the Nickel contract is not respected, then this means that the Nickel representation of the artifact's Dockerfile (Listing 1) does not respect the ``model'' (in the UML sense) / structure and content expected by \texttt{ecg} (the script performing the downloading, the building and the informtation extracting of the Docker image from the research artifact).
-  The transformation from the Nickel representation to the JSON representation (given as input to \texttt{ecg}) will return an error, and the workflow will not even try to rebuild this Docker image.
+  If the Nickel contract is not respected, then this means that the Nickel representation of the artifact's Dockerfile (Listing 1) does not respect the ``model'' (in the UML sense) / structure and content expected by \texttt{ecg} (the script performing the downloading, the building and the information extracting of the Docker image from the research artifact).
+  The transformation from the Nickel representation to the JSON representation (given as input to \texttt{ecg}) (see Figure 1) will return an error, and the workflow will not try to rebuild this Docker image.
+  In short, the contract allows us to verify the correctness of the artifact description (Listing 1) before trying to execute \texttt{ecg}.
 
-  To improve clarity, we swapped Section 2.1 and Section 2.2, and argumented more the use of the Nickel representation.
+  To improve clarity, we swapped Section 2.1 and Section 2.2, and explained more the use of the Nickel representation.
 \end{review-answer}
@@
 \begin{review-answer}
-  This question has been studied in the litterature (see for example the citations [28, 32] ([27, 31] in the initial submission) cited in second paragraph of the introduction), and is not addressed in this paper.
+  This question has been studied in the literature (see for example the citations [28, 32] ([27, 31] in the initial submission) cited in second paragraph of the introduction), and is not addressed in this paper.
 
@@
 \begin{review-answer}
-  While this is true, we do not think that this approach is sustainable.
-  Storing Gigabytes of binary container images on long-term storage is much more expensive than storing a few kilobytes text file.
-  This point is discussed in the Introduction.
+  While this is true, we do not think that this approach is sustainable..
+  Storing Gigabytes of binary container images on long-term storage (\eg\ Zenodo) is much more expensive than storing a few kilobytes text file.
+  This point is discussed in the Introduction and in the Conclusion.
 \end{review-answer}
@@
 The paper proposed by the reviewer does not document this issue: it claims (in the last paragraph of page 31) that there is an issue without showing any proof of it.
-This is a common claim what we have seen in the litterature, but we have yet to see any proof of it.
-Our work therefore aims to bring evidence supporting the claim that the reproducibility of Sofware Environments decays over time.
+This is a common claim what we have seen in the literature, but we have yet to see any proof of it.
+Our work therefore aims to bring evidence supporting the claim that the reproducibility of Software Environments decays over time.
 
-We added a sentence at the end of the introduction to explicit this gap in the litterature.
+We added a sentence at the end of the introduction to explicit this gap in the literature.
 \end{review-answer}
@@
 
-  To improve clarity, we swapped Section 2.1 and Section 2.2, and argumented more the use of the Nickel representation.
+  To improve clarity, we swapped Section 2.1 and Section 2.2, and explained more the use of the Nickel representation.
 \end{review-answer}
@@
 \begin{review-answer}
-  This is correct, and explicited in the \emph{Perspectives} section of the article.
+  This is correct, and stated in the \emph{Perspectives} section of the article.
   As this is a short paper and an initial design of what might be a larger scale experiment, we focused on a conference of our field (HPC), where we more used to the tools and practices.
@@
   \begin{review-answer}
-    While this practice does improve long-term reproductiblity, it lacks scalability and sustainability as storing Gibabytes of binary is much more expensive than storing a few kilobytes of plain text.
-    This point as been discussed in the conclusion section of the article.
+    While this practice does improve long-term reproducibility, it lacks scalability and sustainability as storing Gibabytes of binary is much more expensive than storing a few kilobytes of plain text.
+    This point as been discussed in the introduction and conclusion sections of the article.
   \end{review-answer}
@@
   \begin{review-answer}
-    From our understanding, this only guarentees the top-level dependencies to be correct, but if there are dependencies of those dependeices that have not been captured correctly, the final software environment might still differ.
+    From our understanding, this only guarantees the top-level dependencies to be correct, but if there are dependencies of those dependencies that have not been captured correctly, the final software environment might still differ.
   \end{review-answer}
@@
   \begin{review-answer}
-    Using Git and direct downloads does not guarantees long-term reproducibility.
+    Using Git and direct downloads does not guarantee long-term reproducibility.
     The content behind a direct download can change while the URL stays the same.
@@
 \begin{review-answer}
-  These practices might indeed improve reproducibility, while to garenteeing anything.
-  We agree with the reviewer on the ``rigorously'' and ``improve'' terms.
-  The issue is that it is nearly impossible to create a docker image that will be reproducible over time, but it is very easy to create a docker image that works now.
-  So the reframed question is ``Are container recipes a real solution to the reproducibility problem like the calls for artifacts seem to claim?''.
+  These practices might indeed improve reproducibility, while not bringing strong guarantees.
+  The issue is that it is nearly impossible to create a Docker image that will be reproducible over time, but it is very easy to create a docker image that works now.
+  So the reframed question is ``Are container recipes (like Dockerfiles) a real solution to the reproducibility problem like the calls for artifacts seem to claim?''.
 
@@
 \begin{review-answer}
-  We fully acknowlege that a wider evaluation is required to target a full publication.
+  We fully acknowledge that a wider evaluation is required to target a full publication.
   As stated in the conclusion section (section 5), this short paper aims to be an initial version of a larger scale study.

### 2025-05-20 src/for_reviewers.tex
aspell on the text

@@
 \usepackage{graphicx}
+% \begin{document}
+\title{%
+  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design
+	\\
+	{\Large%
+	Answer to reviewers
+	}
+}
+
+\author{Anonymous Authors}
+\date{}
+
 \begin{document}
 
+\maketitle
+
+\begin{abstract}
+This document is composed of three parts:
+
+\begin{enumerate}
+\item the detailed answers to the reviewers comments
+\item a ``diff'' between the initial submission and the new submission (additions in underlined blue, deletions in crossed-out red)
+\item the new submission of the article
+\end{enumerate}
+\end{abstract}
+
+\tableofcontents
+
+\section{Answers to reviewers}
 \includepdf[pages=-]{rebuttal.pdf}
-\section*{Changes made}
+\section{Changes made}
 \includepdf[pages=-]{diff.pdf}
-\section*{New version}
+\section{New version}
 \includepdf[pages=-]{main.pdf}

### 2025-05-20 src/main.tex
submitted version
aspell on the text

@@
 One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
-Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
+Several works already shed light on the possible variability of experimental results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
 To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
@@
 Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.
+To the best of our knowledge, no previous work has exhibited the variations in software environment build from \df s through time.
 
@@
 
+\subsection{Information captured by \ecg}\label{sec:ecg:capture}
+
+This section presents the different information collected by \ecg.%, and their motivations.
+
+\subsubsection{Cryptographic Hash of the Artifact}
+
+The authors can share artifacts in several ways, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
+Hence, \ecg\ captures the result of the download at the given link (success or error).
+In case of a successful download, \ecg\ computes the cryptographic hash of the download result to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
+
+For example, authors can link to a Zenodo archive that will \emph{always} point to the latest artifact version.
+Hence, the content of the artifact behind this link can change over time.
+
+\subsubsection{Docker Building Error}
+
+In the case of \df\ failing to build, \ecg\ captures the error returned by the building process (\eg\ missing image, failing command).
+In fact, \ecg\ is only capable of capturing the \emph{first error} of the build even if there might be several errors in the building process.
+The Docker error is extracted by \texttt{grep}ing the building process log.
+
+\subsubsection{List of Installed Packages and their Version}
+
+The software environment can be composed of several packages, which can have been installed by different means.
+The most common way is through package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
+But there are also packages installed for programming languages such as with \texttt{pip} for Python.
+Another way to install the package is to download the sources and build them locally.
+\ecg\ is able to capture these different ways to install packages.
+Hence, for each package, \ecg\ extracts the package name, the version, and the tool used to install the package.
+
+Currently, \ecg\ is capable of querying information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are installed by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}), or the case of packages installed in virtual Python environments.
+In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing~\ref{lst:nickel}), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container, and use this commit as a ``version''.
+In the case of a download of the sources of a package (\texttt{misc\_packages} in Listing~\ref{lst:nickel}), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
+The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
+Hence, \ecg\ will download \emph{locally} (outside the container) the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
+
 \subsection{Artifact representation}\label{sec:nickel}
 
-With \ecg, a \df\ from an artifact is represented as a Nickel configuration file~\cite{nickel}.
-Nickel is a configuration language with a functional paradigm.
+The information captured by \ecg\ cannot be automatically extracted from a \df.
+In some cases, authors of artifact can have shell scripts installing packages (\eg\ calling \texttt{apt install}), which would be invisible at the \df\ level.
+With \ecg, a \df\ from an artifact is represented as a Nickel configuration file~\cite{nickel}, which is similar to JSON or YAML.
+%Nickel is a configuration language with a functional paradigm.
 One of Nickel's features is the possibility to \emph{verify} that a configuration file is valid with respect to a \emph{contract}.
@@
 The contract is versioned (line 2 of Listing~\ref{lst:nickel}) to take into account potential future changes to the descriptions and the contract.
+If the Nickel file does not respect the contract, an error is reported to the user.
 
@@
 
-\subsection{Information captured by \ecg}\label{sec:ecg:capture}
-
-This section presents the different information collected by \ecg.%, and their motivations.
-
-\subsubsection{Cryptographic Hash of the Artifact}
-
-The authors can share artifacts in several ways, but not all ways have the same guarantees in terms of longevity and ``stability'' of the link.
-Hence, \ecg\ captures the result of the download at the given link (success or error).
-In case of a successful download, \ecg\ computes the cryptographic hash of the download result to capture any variation in the source of the artifact which could later explain variations in the resulting software environment.
-
-For example, authors can link to a Zenodo archive that will \emph{always} point to the latest artifact version.% of an artifact.
-Hence, the content of the artifact behind this link can change over time.
-
-\subsubsection{Docker Building Error}
-
-In the case of \df\ failing to build, \ecg\ captures the error returned by the building process (\eg\ missing image, failing command).
-In fact, \ecg\ is only capable of capturing the \emph{first error} of the build even if there might be several errors in the building process.
-The Docker error is extracted by \texttt{grep}ing the building process log.
-
-\subsubsection{List of Installed Packages and their Version}
-
-The software environment can be composed of several packages, which can have been installed by different means.
-The most common way is through package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
-But there are also packages installed for programming languages such as with \texttt{pip} for Python.
-Another way to install the package is to download the sources and build them locally.
-\ecg\ is able to capture these different ways to install packages.
-Hence, for each package, \ecg\ extracts the package name, the version, and the tool used to install the package.
-
-Currently, \ecg\ is capable of querying information from popular package managers (\texttt{dkpg} (and thus \texttt{apt}), \texttt{pacman}, \texttt{pip}, \texttt{conda}) and to deal with situations where packages are installed by source (\texttt{package\_managers} in Listing~\ref{lst:nickel}), or the case of packages installed in virtual Python environments.
-In the case where there is a \texttt{git clone} command in the \df\ (\texttt{git\_packages} in Listing~\ref{lst:nickel}), \ecg\ will extract the commit of the local clone of the repository after the successful build of the container, and use this commit as a ``version''.
-In the case of a download of the sources of a package (\texttt{misc\_packages} in Listing~\ref{lst:nickel}), for example with \texttt{wget} or \texttt{curl}, there is no immediate way to log the version of the package that is being downloaded.
-The URL of the source code might contain a version number, but this URL might in the future point to a different version of the source code.
-Hence, \ecg\ will download \emph{locally} (outside the container) the source code pointed by the URL and compute the cryptographic hash of the result to generate a ``version''.
-
 \subsection{Workflow}\label{sec:workflow}
@@
 The workflow of \ecg\ is shown in Figure~\ref{fig:workflow}.
-The workflow itself~\cite{ecg_workflow} is governed by a Snakemake~\cite{koster2012snakemake} description, and its software environment~\cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.
+The workflow itself~\cite{ecg_workflow} is governed by the Snakemake workflow manager~\cite{koster2012snakemake}, and its software environment~\cite{ecg_nix} is managed by the Nix functional package manager~\cite{dolstra_nix_2004}.
+These tools provide reproducibility and robustness to the workflow setup to ensure that we will be able to execute the exact same workflow in the exact same conditions for all the future builds of the artifacts.
 
@@
 \item For each conference with an artifact evaluation process, we examine all artifacts and consider only those that share their software environment using a \df.
-\item From the artifact and the \df we generate the associated Nickel file (Listing~\ref{lst:nickel}) which contains information such as the artifact's location, how to extract the \df\ from the artifact and the package managers used.
+\item From the artifact and the \df\ we generate the associated Nickel file (Listing~\ref{lst:nickel}) which contains information such as the artifact's location, how to extract the \df\ from the artifact and the package managers used.
 %where is the artifact stored? how to extract the \df\ out from the artifact? which package managers are used?
@@
 \item \ecg\ is executed on the x86-64 machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
+\item We do not try to reproduce the experiment results from the artifacts, but only to build the container.
+The impact of software environment variations on the experiments results has been studied for example in \cite{mytkowicz_producing_nodate, sokolowski2024impact}.
 \end{itemize}
@@
 
-The community may want to adopt solutions \emph{tailored for reproducibility} of software environment (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual representation of their software environment which could be precisely reproduced and with longevity guarantees~\cite{courtes2024source}.
+The community may want to adopt instead solutions \emph{tailored for long-term reproducibility} of software environment~\cite{malka2024reproducibility} (\ie\ Nix~\cite{dolstra_nix_2004} and Guix~\cite{courtes_functional_2013}), where authors can share a textual representation of their software environment which could be precisely reproduced and with longevity guarantees~\cite{courtes2024source}.
 Moreover, Nix and Guix allow the precise introduction of variation in the software environment, which is a crucial feature (not present in container technologies) in order to build upon previous artifact for future publications~\cite{mercier2018considering}.
@@
 
+\newpage
 \section*{Acknowledgments}
@@
 
-\newpage
-\newpage
+% \newpage
+% \newpage
 %\bibliographystyle{sty/ACM-Reference-Format}

### 2025-05-21 src/rebuttal.tex
rebuttal review from florina

@@
 
-\usepackage[%
+\usepackage[
 	a4paper,
@@
 
-% \maketitle
+%\maketitle
 
@@
   This initial study is indeed limited to five artifact from a single conference (Euro-Par24).
-  The goal of this short paper is to investigate the necessity and potential of a larger scale study.
+  The purpose of this short paper is to investigate the necessity and potential of a larger-scale study.
 \end{review-answer}
@@
   We cannot simply use/parse Dockerfiles to extract the same information.
-  Indeed, the authors crafting their Dockerfile could for example do the installation of their software dependencies in a shell script that will then be called in the steps of the Dockerfile.
-  The content of this script is unknown at the level of the Dockerfile.
-  Hence the need to capture manually the download and installations, and list them in this Nickel format.
+  Indeed, the authors crafting their Dockerfile could, for example, do the installation of their software dependencies in a shell script that will then be called in the steps of the Dockerfile.
+  The content of this script is unknown at the Dockerfile level.
+  Hence, the need to capture manually the download and installations and list them in this Nickel format.
 
-  To improve clarity, we swapped Section 2.1 and Section 2.2, and explained more the use of the Nickel representation.
+  To improve clarity, we swapped the order of Sections 2.1 and 2.2, and explained more about the use of the Nickel representation.
 \end{review-answer}
@@
       \begin{review-answer}
-        We added a description of those tools in Section 2.3.
+        Thank you - we now added a description of these tools in Section 2.3.
       \end{review-answer}
@@
         \begin{itemize}
-        \item The impact of software environment variations on experiments results have been heavily studied (see references [28, 32] ([27, 31] in the initial submission), cited in the second paragraph of the introduction)
-        \item The energy-costs of actually reproducing the experiments for which the toolchain determined that the environment can be built must be weighed against the utility of reproducing those experiments.
-          The goal of our short paper is to prove the concept of experiment building over time.
-          This benefits researchers that indeed want/need to reproduce specific experiments and justify the associated energy-costs
-        \end{itemize}
-
-        We added a sentence in Section 3 to make this point more explicit.
+        \item The impact of software environment variations on experiments results has been widely studied (see references [28, 32] ([27, 31] in the initial submission), cited in the second paragraph of the introduction.
+        \item The energy costs of actually reproducing the experiments for which our proposed toolchain determined that the environment can be built must be weighed against the utility of reproducing those experiments.
+          The goal of our short paper is to prove the \textit{concept} of experiment building over time.
+          This benefits researchers who in fact want/need to reproduce specific experiments and can then justify the associated energy costs.
+        \end{itemize}.
+        In Section 3 we added a sentence to make this point more explicit.
       \end{review-answer}
@@
 \begin{review-answer}
-  Indeed, the approach needs to be further developed and tested with a larger collection of artifacts, as stated in the \emph{Perspectives} section.
-  The extent of our current submission is a short-paper describing a proof of concept.
+  We agree that the approach needs to be further developed and tested with a larger collection of artifacts, as stated in the \emph{Perspectives} section.
+  The scope of our current submission is a short paper that describes a proof of concept that experiments can be built over time.
 \end{review-answer}
@@
 
-\subsection{Strenght}
+\subsection{Strengths}
 
@@
 
-  If the Nickel contract is not respected, then this means that the Nickel representation of the artifact's Dockerfile (Listing 1) does not respect the ``model'' (in the UML sense) / structure and content expected by \texttt{ecg} (the script performing the downloading, the building and the information extracting of the Docker image from the research artifact).
-  The transformation from the Nickel representation to the JSON representation (given as input to \texttt{ecg}) (see Figure 1) will return an error, and the workflow will not try to rebuild this Docker image.
-  In short, the contract allows us to verify the correctness of the artifact description (Listing 1) before trying to execute \texttt{ecg}.
+  If the Nickel contract is not respected, then the Nickel representation of the artifact's Dockerfile (Listing 1) does not respect the ``model'' (in the UML sense) / structure and content expected by \texttt{ecg} (the script performing the downloading, the building and the information extracting of the Docker image from the research artifact).
+  Transformation from the Nickel representation to the JSON representation (given as input to \texttt{ecg}) (see Figure 1) will return an error and the workflow will not try to rebuild this Docker image.
+  In short, the contract allows us to verify the correctness of the artifact description (Listing 1) before attempting to execute \texttt{ecg}.
 
-  To improve clarity, we swapped Section 2.1 and Section 2.2, and explained more the use of the Nickel representation.
+  To improve clarity, we swapped the order of Sections 2.1 and 2.2, and explained more about the use of the Nickel representation.
 \end{review-answer}
@@
   \begin{review-answer}
-    We added a citation this paper in the Conclusion section (Reference [25]).
+    Thanks. We now cited this paper in the Conclusion section (Reference [25]).
   \end{review-answer}
@@
 \begin{review-answer}
-  This question has been studied in the literature (see for example the citations [28, 32] ([27, 31] in the initial submission) cited in second paragraph of the introduction), and is not addressed in this paper.
+  This question has been studied in the literature (see, for example, the citations [28, 32] ([27, 31] in the initial submission) cited in the second paragraph of the introduction and is not addressed in this paper.
 
-We added a sentence in Section 3 to make this point more explicit.
+In Section 3 we added a sentence to make this point more explicit.
 \end{review-answer}
+
 An easy fix would be to remove the apt-update and keeping the base image with a version.
 \begin{review-answer}
-  The preliminary results presented in this paper show that using tools such as \texttt{pip} also introduce variation in their produced software environment (see Figure 3).
+  The preliminary results presented in this paper show that the use of tools such as \texttt{pip} also introduce variation in their produced software environment (see Figure 3).
 \end{review-answer}
@@
 \begin{review-answer}
-  While this is true, we do not think that this approach is sustainable..
-  Storing Gigabytes of binary container images on long-term storage (\eg\ Zenodo) is much more expensive than storing a few kilobytes text file.
+  While this is true, we do not believe that this approach is sustainable.
+  Storing Gigabytes of binary container images on long-term storage (\eg Zenodo) is much more expensive than storing a few kilobyte text file.
   This point is discussed in the Introduction and in the Conclusion.
@@
 \begin{review-answer}
-  The typos have been corrected.
+  Thanks. The typos have now been corrected.
 \end{review-answer}
@@
 \begin{review-answer}
-The paper proposed by the reviewer does not document this issue: it claims (in the last paragraph of page 31) that there is an issue without showing any proof of it.
-This is a common claim what we have seen in the literature, but we have yet to see any proof of it.
+The paper proposed by the reviewer does not document this issue: it claims (last paragraph, page 31) that there is an issue without showing any proof of it.
+This is a common claim of the literature, but we are unaware of any proof or evidence supporting it. %have yet to see any proof of it.
 Our work therefore aims to bring evidence supporting the claim that the reproducibility of Software Environments decays over time.
 
-We added a sentence at the end of the introduction to explicit this gap in the literature.
+We added a sentence at the end of the introduction to make this gap explicit in the literature.
 \end{review-answer}
@@
   There might be a misunderstanding.
-  We do not propose another format to generate and share container images.
-  We use the textual representation presented in Section 2.2 (Section 2.1 in the initial submission) to track what are the downloads and installations done in a Dockerfile from a research artifact.
-  We then build the Docker image from the Dockerfile, and use our representation to perform the operation required to extract information about the generated software environment.
+  We do not propose another format for generating and sharing container images.
+  We use the textual representation presented in Section 2.2 (Section 2.1 in the initial submission) to track the downloads and installations performed in a Dockerfile from a research artifact.
+  We then build the Docker image from the Dockerfile and use our representation to perform the operation required to extract information about the generated software environment.
   We need this representation because we cannot simply use the Dockerfile to extract the information about which package is installed and how.
@@
 
-  To improve clarity, we swapped Section 2.1 and Section 2.2, and explained more the use of the Nickel representation.
+  To improve clarity, we swapped the order of Sections 2.1 and 2.2, and explained more about the use of the Nickel representation.
 \end{review-answer}
@@
 \begin{review-answer}
-  This is correct, and stated in the \emph{Perspectives} section of the article.
-  As this is a short paper and an initial design of what might be a larger scale experiment, we focused on a conference of our field (HPC), where we more used to the tools and practices.
-  Previous study have shown that Docker is the main container tool used to produce container for research artifacts in HPC (see reference [20]), hence this initial focus on Docker.
+  This is correct and is also stated in the \emph{Perspectives} section of the article.
+  As this is a short paper that presents an initial design of what might be a larger-scale experiment, we focus on a conference of our field (HPC) where we are more familiar with the tools and practices.
+  Previous studies have shown that Docker is the main container tool used to produce containers for research artifacts in HPC (see reference [20]), hence supporting Docker as an initial focus in this work.
 \end{review-answer}
@@
 \begin{review-answer}
-  There are indeed best practices than artifact authors can follow.
-  Let us note that the practices of academia differ to the practices of industry, where the objectives are different.
-  Having authors following better practices would be ideal, but unfortunately it is not the case.
+  Indeed, there are best practices that artifact authors can follow.
+  We note that the practices of academia differ from the practices of industry, where the objectives are different.
+  Having authors following better practices would be ideal, but unfortunately it is not the case and also not only up to us (authors of this short paper).
 \end{review-answer}
@@
     \begin{review-answer} 
-    While this practice does improve long-term reproducibility, it lacks scalability and sustainability as storing Gibabytes of binary is much more expensive than storing a few kilobytes of plain text.
-    This point as been discussed in the introduction and conclusion sections of the article.
+    While this practice does improve long-term reproducibility, it lacks scalability and sustainability, as storing Gibabytes of binary is much more expensive than storing a few kilobytes of plain text.
+    This point has been discussed in the introduction and conclusion sections of the article.
   \end{review-answer}
@@
     The content behind a direct download can change while the URL stays the same.
-    Git commits can be remove/rewritten and repositories deleted.
+    Git commits can be removed/rewritten and repositories deleted.
 
-    This point has been discussed in the sections 2.1.1 and 2.1.3 (sections 2.2.1 and 2.2.3 in the initial submission).
+    This point has been discussed in current Sections 2.1.1 and 2.1.3 (old Sections 2.2.1 and 2.2.3 in the initial submission).
   \end{review-answer}
@@
   The issue is that it is nearly impossible to create a Docker image that will be reproducible over time, but it is very easy to create a docker image that works now.
-  So the reframed question is ``Are container recipes (like Dockerfiles) a real solution to the reproducibility problem like the calls for artifacts seem to claim?''.
+  Therefore, the reframed question is ``Are container recipes (like Dockerfiles) a real solution to the reproducibility problem like the calls for artifacts seem to claim?''.
 
@@
 \begin{review-answer}
-  We are not proposing a new format to build and share container images.
+  We are not proposing a new format for building and sharing container images.
 \end{review-answer}
@@
   We fully acknowledge that a wider evaluation is required to target a full publication.
-  As stated in the conclusion section (section 5), this short paper aims to be an initial version of a larger scale study.
+  As stated in the conclusion section (Section 5), this short paper aims to be an initial version of a larger-scale study.
 \end{review-answer}

### 2025-05-24 src/for_reviewers.tex
with the 1-page rebuttal

--- 2025-05-19 src/for_reviewers.tex
@@
 \usepackage{graphicx}
+\usepackage[
+	a4paper,
+	margin=20mm,  % tighter margins
+]{geometry}
+\pagenumbering{gobble}
+% \begin{document}
+\title{%
+  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design
+	\\
+	{\Large%
+	Answer to reviewers
+	}
+}
+
+\author{Anonymous Authors}
+\date{}
+
 \begin{document}
 
+\maketitle
+
+% \begin{abstract}
+% This document is composed of three parts:
+% 
+% \begin{enumerate}
+% \item a 1-page answer addressing comments from the reviewers
+% \item the new submission of the article
+% \item the detailed answers to the reviewers comments
+% \item a ``diff'' between the initial submission and the new submission (additions in underlined blue, deletions in crossed-out red)
+% \end{enumerate}
+% \end{abstract}
+
+% \tableofcontents
+
+First and foremost, we would like to thank the reviewers for their insightful comments.
+In this 1-page document, we address their recurrent interrogations.
+Following this 1-page document, we attached the revised version of the article.
+For completeness, we also attached detailed and individual answers to the reviewers' comments, as well as a clear visual representation of the changes between the initial submission and the revised version of the article (additions are underlined in blue, and deletions in crossed-out red).
+
+\paragraph{Usage of Nickel}
+
+Nickel is a configuration language (like JSON or YAML), but with the possibility to define constraints on the configurations' structures and values, named ``contract''s.
+
+We cannot simply use/parse Dockerfiles to extract the same information.
+Indeed, the authors crafting their Dockerfile could, for example, do the installation of their software dependencies in a shell script that will then be called in the steps of the Dockerfile.
+The content of this script is unknown at the Dockerfile level.
+Hence, the need to capture manually the download and installations and list them in this textual Nickel format.
+
+If the Nickel contract is not respected, then the Nickel representation of the artifact's Dockerfile (Listing 1) does not respect the ``model'' (in the UML sense) / structure and content expected by \texttt{ecg} (the script performing the downloading, the building and the information extracting of the Docker image from the research artifact).
+Transformation from the Nickel representation to the JSON representation (given as input to \texttt{ecg}) (see Figure 1) will return an error and the workflow will not try to rebuild this Docker image.
+In short, the contract allows us to verify the correctness of the artifact description (Listing 1) before attempting to execute \texttt{ecg}.
+
+This Nickel representation is therefore not an attempt at replacing Dockerfile, but is instead of extracted view of what the Dockerfile will download and install, such that the information about what has actually been downloaded and installed can be extracted by \texttt{ecg}.
+This representation is in no way a proposition for a new tool/format to build container images.
+
+$\hookrightarrow$ \textbf{To improve clarity, we swapped the order of Sections 2.1 and 2.2, and explained more about the use of the Nickel representation.}
+
+\paragraph{Impact of software environment on experiments results}
+
+The question of the impact of variation software environment on the experiments results has been studied in the literature (see, for example, the citations [28, 32] ([27, 31] in the initial submission) cited in the second paragraph of the introduction and is not addressed in this paper.
+In this work, we thus do not try to reproduce the experiments associated to the Dockerfile that we are building.
+
+$\hookrightarrow$ \textbf{In Section 3 we added a sentence to make this point more explicit.}
+
+\paragraph{Limited evaluation}
+
+We fully acknowledge that a wider evaluation is required to target a full publication.
+As stated in the conclusion section (Section 5), this short paper aims to be an initial version of a larger-scale study.
+The scope of our current submission is a short paper that describes a \emph{proof of concept} that experiments can be built over time.
+The purpose of this short paper is to investigate the necessity and potential of a larger-scale study.
+This wider evaluation will need to consider several container technologies, from various conferences with varying levels of computer science technical expertise.
+
+While the drifting of the software environment produced by Dockerfile has been a common claim of the literature, we are unaware of any proof or evidence supporting it and characterizing it.
+%have yet to see any proof of it.
+Our work therefore aims to bring evidence supporting the claim that the reproducibility of Software Environments decays over time.
+
+$\hookrightarrow$ \textbf{We added a sentence at the end of the introduction to make this gap explicit in the literature.}
+
+\newpage
+% \section{New version}
+\includepdf[pages=-]{main.pdf}
+% \section{Answers to reviewers}
 \includepdf[pages=-]{rebuttal.pdf}
-\section*{Changes made}
+% \section{Changes made}
 \includepdf[pages=-]{diff.pdf}
-\section*{New version}
-\includepdf[pages=-]{main.pdf}
 \end{document}

### 2025-06-09 src/main.tex
with the 1-page rebuttal
de-anonymise

@@
 \title{%
-  % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
   SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
@@
 
-% \author{Quentin Guilloteau}
-% \orcid{0009-0003-7645-5044}
-% \email{Quentin.Guilloteau@inria.fr}
+\author{Quentin Guilloteau}
+\orcid{0009-0003-7645-5044}
+\email{Quentin.Guilloteau@inria.fr}
+\affiliation{%
+  \institution{Inria, CNRS, ENSL, UCBL1, LIP}
+  \city{Lyon}
+  \country{France}
+}
+
+\author{Antoine Waehren}
+\orcid{0009-0003-9852-8865}
+\email{A.Waehren@stud.unibas.ch}
+\affiliation{%
+  \institution{University of Basel}
+  \city{Basel}
+  \country{Switzerland}
+}
+
+\author{Florina M. Ciorba}
+\orcid{0000-0002-2773-4499}
+\email{Florina.Ciorba@unibas.ch}
+\affiliation{%
+  \institution{University of Basel}
+  \city{Basel}
+  \country{Switzerland}
+}
+
+% \author{Anonymous Authors}
+% \email{Anonymous email}
 % \affiliation{%
-%   \institution{Inria, CNRS, ENSL, UCBL1, LIP}
-%   \city{Lyon}
-%   \country{France}
+%   \institution{Anonymous Institution}
+%   \city{Anonymous City}
+%   \country{Anonymous Country}
 % }
-% 
-% \author{Antoine Waehren}
-% \email{A.Waehren@stud.unibas.ch}
-% \affiliation{%
-%   \institution{University of Basel}
-%   \city{Basel}
-%   \country{Switzerland}
-% }
-% 
-% \author{Florina M. Ciorba}
-% \email{Florina.Ciorba@unibas.ch}
-% \affiliation{%
-%   \institution{University of Basel}
-%   \city{Basel}
-%   \country{Switzerland}
-% }
-
-\author{Anonymous Authors}
-\email{Anonymous email}
-\affiliation{%
-  \institution{Anonymous Institution}
-  \city{Anonymous City}
-  \country{Anonymous Country}
-}
 
@@
 The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
-The deadline for the authors to submit their artifacts to Euro-Par 2024 was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
+The deadline for the authors to submit their artifacts to Euro-Par 2024 was May 15$^{th}$ 2024.
+%\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
 Hence, our initial build took place almost 5 months after the artifacts submission.
